### 2020-08-03 notebooks/Eigenvector_investigation_0528.Rmd
added 2 notebooks

---
itle: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4,fig.height = 4)
```

```{r}
library(tidyverse)
```

```{r}
eigvecs <- data.table::fread('../data/interim/eigvecs.txt')%>%rownames_to_column('id')%>%mutate(id=as.numeric(id))
eigvals <- data.table::fread('../data/interim/eigvals.txt')
#getwd()

eig_l0 <- eigvecs[,2]#%>%pull()
eig_l1116 <- eigvecs[,1118]#%>%pull()
```

```{r}
ggplot(eigvecs,aes(x=id,y=sort(V1)))+geom_point()+geom_hline(aes(yintercept=0,color='red'))

ggplot(eigvecs,aes(x=id,y=sort(V1117)))+geom_point()+geom_hline(aes(yintercept=0,color='red'))
```

```{r}
summary(eig_l0)
summary(eig_l1116)
```

```{r}
library(igraph)
g_file = read_csv('../data/interim/costanzo_pcc_ALL')
g = graph_from_data_frame(g_file,directed = FALSE)%>%igraph::simplify()
gc = induced_subgraph(g,V(g)[components(g)$membership==1])
# ee = embed_laplacian_matrix(gc,5182)
# L = laplacian_matrix(gc)
enm_df = read_csv('../data/interim/enm_df.csv')
tictoc::tic()
lost.edges <- c()
for(i in 1:(vcount(gc)-1)){#seq_along(V(gc))){
  if (i ==vcount(gc)){
    next
  }
   c1 = induced.subgraph(gc,V(gc)[enm_df$orf_name[eigvecs[,i+1]>0]])
   c2= induced.subgraph(gc,V(gc)[enm_df$orf_name[eigvecs[,i+1]<0]])
   lost.edges <- c(lost.edges, (ecount(gc)-ecount(c1)-ecount(c2)))
  print(i,length(lost.edges))
   }
tictoc::toc()
plot(lost.edges)
```