### 2020-11-11 notebooks/paper_plots_0909.Rmd
paper plots include a combined plot of density distributions.

---
title: "Paper figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
#use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv('enm')
setwd('../')
```

```{python}
import matplotlib.pyplot as plt
import matplotlib as mpl
import networkx as nx
import numpy as np
import pandas as pd
import plotnine as p9
import pickle 
import os
import re
import itertools as itr
from src.enm import Enm

with open(f'data/interim/pcc_0909/pcc.pickle','rb') as f:
    e_pcc = pickle.load(f)
with open(f'data/interim/yuri_0916/yuri.pickle','rb') as f:
    e_yuri = pickle.load(f)
with open(f'data/interim/huri_0916/huri.pickle','rb') as f:
    e_huri = pickle.load(f)

pcc_df = e_pcc.df
pcc_gc = e_pcc.graph_gc
pcc_gc_pos = (dict(pcc_gc.nodes(data='pos')))
yuri_df = e_yuri.df
huri_df = e_huri.df
```

```{r}
library(tidyverse)
library(cowplot)
library(ggpubr)

pcc_df_random  = data.table::rbindlist(lapply(py$e_pcc$e_list, function(x) x$df),fill=T,idcol='random_id')%>%mutate(random_id = as.factor(random_id))
yuri_df_random  = data.table::rbindlist(lapply(py$e_yuri$e_list, function(x) x$df),fill=T,idcol='random_id')%>%mutate(random_id = as.factor(random_id))
huri_df_random  = data.table::rbindlist(lapply(py$e_huri$e_list, function(x) x$df),fill=T,idcol='random_id')%>%mutate(random_id = as.factor(random_id))
pcc_df = py$pcc_df%>%mutate(data='pcc')%>%mutate(sens_norm = sens/nrow(.),eff_norm = eff/nrow(.), group = 'real')%>%bind_rows(
    pcc_df_random%>%mutate(data='pcc')%>%mutate(sens_norm = sens/nrow(py$pcc_df),eff_norm = eff/nrow(py$pcc_df), group = 'random')
)%>%mutate(group=fct_relevel(group, c('real','random')))
yuri_df = py$yuri_df%>%mutate(data='yuri')%>%mutate(sens_norm = sens/nrow(.),eff_norm = eff/nrow(.), group = 'real')%>%bind_rows(
    yuri_df_random%>%mutate(data='yuri')%>%mutate(sens_norm = sens/nrow(py$yuri_df),eff_norm = eff/nrow(py$yuri_df), group = 'random')
)%>%mutate(group=fct_relevel(group, c('real','random')))
huri_df = py$huri_df%>%mutate(data='huri')%>%mutate(sens_norm = sens/nrow(.),eff_norm = eff/nrow(.), group = 'real')%>%bind_rows(
    huri_df_random%>%mutate(data='huri')%>%mutate(sens_norm = sens/nrow(py$huri_df),eff_norm = eff/nrow(py$huri_df), group = 'random')
)%>%mutate(group=fct_relevel(group, c('real','random')))
combined_df = bind_rows(pcc_df, yuri_df,huri_df)%>%
    mutate(data= fct_relevel(data, c('pcc','yuri','huri')))
```

Degree - Eff - Sens distributions
```{r}
combined_df%>%filter(group=='real')%>%
    ggplot(aes(x=deg, y = ..count../sum(..count..)))+
    facet_grid(rows=vars(data))+
    geom_histogram(binwidth=1)+
#    xlim(1,NA)+
    ylim(NA, 0.1)

pcc_degree_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data))+
    #geom_density()+
    geom_line(data=     pcc_df%>%filter(group=='real'),aes(x=deg,y=..scaled..,color=group),stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data=     pcc_df%>%filter(group=='random'),aes(x=deg,y=..scaled.., color=group, fill=random_id),stat = 'density',linetype = "dashed")+#, binwidth = 0.1) + 
    theme_bw()+
    xlab('Degree')+
    ylab('Density')+
    coord_trans(x = "log10")+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
theme(panel.grid.minor = element_blank())#+theme_fontsize
pcc_eff_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data))+
#    geom_density()+
    geom_line(data= pcc_df%>%filter(group=='real'),
               aes(x=eff_norm ,y=..scaled.., color=group),
               stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data= pcc_df%>%filter(group=='random'),
               aes(x=eff_norm ,y=..scaled..,fill=random_id, color=group),
               stat = 'density',linetype = "dashed" )+#, binwidth = 0.1) + 
    #geom_histogram(binwidth=0.0001)+
    theme_bw()+
    ylab('Density')+
    xlab('Effectiveness')+
    coord_trans(x = "log10")+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
theme(panel.grid.minor = element_blank())#+theme_fontsize+
#annotate("text", x = 0.000001, y = 0.2, label=sprintf("p = %#.2f", eff_p), fontface='italic', size = 6)
pcc_sensitivity_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data), scales='free_y')+
#    geom_density()+
    geom_line(data= pcc_df%>%filter(group=='real'),
               aes(x=sens_norm,y=..scaled.. , color=group),
               stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data= pcc_df%>%filter(group=='random'),
               aes(x=sens_norm,y=..scaled..,fill=random_id,color=group),
               stat = 'density',linetype = "dashed")+#, binwidth = 0.1) + 
    #geom_histogram(binwidth=0.0001)+
    theme_bw()+
    xlab('Sensitivity')+
    ylab('Density')+
    coord_trans(x = "log10")+
    scale_color_discrete(labels=c('Rewired networks','Real network'))+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 )+ 
# scale_y_continuous(
#   breaks = scales::trans_breaks("log1p", function(x) 10^(1+x)),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
theme(panel.grid.minor = element_blank(), legend.position='bottom', legend.title=element_blank())
#theme_fontsize+
#annotate("text", x = 0.001, y = 0.9, label=sprintf("p = %#.2f", sens_p), fontface='italic', size=6)
title <- cowplot::ggdraw() + cowplot::draw_label("Costanzo 2016 network", fontface='bold')
pcc_dens_comb = cowplot::plot_grid(pcc_degree_dens+theme(legend.position='none'), 
                                pcc_eff_dens+theme(legend.position='none'), 
                                pcc_sensitivity_dens+theme(legend.position='none'), labels=c('A','B','C'),ncol=3)
pcc_dens_comb_title = cowplot::plot_grid(title, pcc_dens_comb, nrow=2 ,rel_heights=c(0.1,0.9))

yuri_degree_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data))+
    #geom_density()+
    geom_line(data=     yuri_df%>%filter(group=='real'),aes(x=deg,y=..scaled..,color=group),stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data=     yuri_df%>%filter(group=='random'),aes(x=deg,y=..scaled.., color=group, fill=random_id),stat = 'density',linetype = "dashed")+#, binwidth = 0.1) + 
    theme_bw()+
    xlab('Degree')+
    ylab('Density')+
    coord_trans(x = "log10")+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
theme(panel.grid.minor = element_blank())#+theme_fontsize
yuri_eff_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data))+
#    geom_density()+
    geom_line(data= yuri_df%>%filter(group=='real'),
               aes(x=eff_norm ,y=..scaled.., color=group),
               stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data= yuri_df%>%filter(group=='random'),
               aes(x=eff_norm ,y=..scaled..,fill=random_id, color=group),
               stat = 'density',linetype = "dashed" )+#, binwidth = 0.1) + 
    #geom_histogram(binwidth=0.0001)+
    theme_bw()+
    ylab('Density')+
    xlab('Effectiveness')+
    coord_trans(x = "log10")+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
theme(panel.grid.minor = element_blank())#+theme_fontsize+
#annotate("text", x = 0.000001, y = 0.2, label=sprintf("p = %#.2f", eff_p), fontface='italic', size = 6)
yuri_sensitivity_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data), scales='free_y')+
#    geom_density()+
    geom_line(data= yuri_df%>%filter(group=='real'),
               aes(x=sens_norm,y=..scaled.. , color=group),
               stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data= yuri_df%>%filter(group=='random'),
               aes(x=sens_norm,y=..scaled..,fill=random_id,color=group),
               stat = 'density',linetype = "dashed")+#, binwidth = 0.1) + 
    #geom_histogram(binwidth=0.0001)+
    theme_bw()+
    xlab('Sensitivity')+
    ylab('Density')+
    coord_trans(x = "log10")+
    scale_color_discrete(labels=c('Rewired networks','Real network'))+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 )+ 
# scale_y_continuous(
#   breaks = scales::trans_breaks("log1p", function(x) 10^(1+x)),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
theme(panel.grid.minor = element_blank(), legend.position='bottom', legend.title=element_blank())
title <- cowplot::ggdraw() + cowplot::draw_label("Yeast PPI network", fontface='bold')
yuri_dens_comb = cowplot::plot_grid(yuri_degree_dens+theme(legend.position='none'), 
                                yuri_eff_dens+theme(legend.position='none'), 
                                yuri_sensitivity_dens+theme(legend.position='none'), labels=c('D','E','F'),ncol=3)
yuri_dens_comb_title = cowplot::plot_grid(title, yuri_dens_comb, nrow=2 ,rel_heights=c(0.1,0.9))

huri_degree_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data))+
    #geom_density()+
    geom_line(data=     huri_df%>%filter(group=='real'),aes(x=deg,y=..scaled..,color=group),stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data=     huri_df%>%filter(group=='random'),aes(x=deg,y=..scaled.., color=group, fill=random_id),stat = 'density',linetype = "dashed")+#, binwidth = 0.1) + 
    theme_bw()+
    xlab('Degree')+
    ylab('Density')+
    coord_trans(x = "log10")+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
theme(panel.grid.minor = element_blank())#+theme_fontsize
huri_eff_dens <- 
    ggplot()+
    #facet_grid(rows=vars(data))+
#    geom_density()+
    geom_line(data= huri_df%>%filter(group=='real'),
               aes(x=eff_norm ,y=..scaled.., color=group),
               stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data= huri_df%>%filter(group=='random'),
               aes(x=eff_norm ,y=..scaled..,fill=random_id, color=group),
               stat = 'density',linetype = "dashed" )+#, binwidth = 0.1) + 
    #geom_histogram(binwidth=0.0001)+
    theme_bw()+
    ylab('Density')+
    xlab('Effectiveness')+
    coord_trans(x = "log10")+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
theme(panel.grid.minor = element_blank())#+theme_fontsize+
#annotate("text", x = 0.000001, y = 0.2, label=sprintf("p = %#.2f", eff_p), fontface='italic', size = 6)
huri_sensitivity_dens <- 
    ggplot()+
#$    facet_grid(rows=vars(data), scales='free_y')+
#    geom_density()+
    geom_line(data= huri_df%>%filter(group=='real'),
               aes(x=sens_norm,y=..scaled.. , color=group),
               stat = 'density')+#, binwidth = 0.1) + 
    geom_line(data= huri_df%>%filter(group=='random'),
               aes(x=sens_norm,y=..scaled..,fill=random_id,color=group),
               stat = 'density',linetype = "dashed")+#, binwidth = 0.1) + 
    #geom_histogram(binwidth=0.0001)+
    theme_bw()+
    xlab('Sensitivity')+
    ylab('Density')+
    coord_trans(x = "log10")+
    scale_color_discrete(labels=c('Rewired networks','Real network'))+
scale_x_continuous(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 )+ 
# scale_y_continuous(
#   breaks = scales::trans_breaks("log1p", function(x) 10^(1+x)),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
theme(panel.grid.minor = element_blank(), legend.position='bottom', legend.title=element_blank())
title <- cowplot::ggdraw() + cowplot::draw_label("HuRI network", fontface='bold')
huri_dens_comb = cowplot::plot_grid(huri_degree_dens+theme(legend.position='none'), 
                                huri_eff_dens+theme(legend.position='none'), 
                                huri_sensitivity_dens+theme(legend.position='none'), labels=c('G','H','I'),ncol=3)
huri_dens_comb_title = cowplot::plot_grid(title, huri_dens_comb, nrow=2 ,rel_heights=c(0.1,0.9))
dens_lgn = cowplot::get_legend(pcc_sensitivity_dens)
dens_plots = cowplot::plot_grid(pcc_dens_comb_title,
                                yuri_dens_comb_title,
                                huri_dens_comb_title,nrow=3)
cowplot::plot_grid(dens_plots, dens_lgn,nrow=2,rel_heights=c(.9,.1))
ggsave('reports/figures/paper_figures_0916/fig1.png')

#degree_dens <- 
#    ggplot(combined_df%>%filter(group=='real'), aes(x=deg))+
#    ggplot(combined_df%>%filter(group=='random'), aes(x=deg, y=..scaled.., fill=.id))+
#    facet_grid(rows=vars(data))+
#    #geom_density()+
#    geom_point(stat = 'density')+#, binwidth = 0.1) + 
#    theme_bw()+
#    xlab('Degree')+
#    coord_trans(x = "log10", y = "log1p")+
#scale_x_continuous(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
# scale_y_continuous(
#   breaks = scales::trans_breaks("log10", function(x) 10^(x)),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
#theme(panel.grid.minor = element_blank())
#eff_dens <- 
#    combined_df%>%filter(group=='real')%>%
#    ggplot(aes(x=eff_norm ))+
#    facet_grid(rows=vars(data), scales='free_y')+
##    geom_density()+
#    geom_point(stat = 'density')+#, binwidth = 0.1) + 
#    #geom_histogram(binwidth=0.0001)+
#    theme_bw()+
#    xlab('Effectivity')+
#    coord_trans(x = "log10", y = "log1p")+
#scale_x_continuous(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
# scale_y_continuous(
#   breaks = scales::trans_breaks("log1p", function(x) 10^(1+x)),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
#theme(panel.grid.minor = element_blank())
#sensitivity_dens <- 
#    combined_df%>%filter(group=='real')%>%
#    ggplot(aes(x=sens_norm))+
#    facet_grid(rows=vars(data))+
#    geom_point(stat = 'density')+#, binwidth = 0.1) + 
#   # geom_density()+
#    theme_bw()+
#    xlab('Sensitivity')+
#    coord_trans(x = "log10", y = "log1p")+
#   # annotation_logticks()+
#scale_x_continuous(
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
# scale_y_continuous(
#   breaks = scales::trans_breaks("log1p", function(x) 10^(1+x)),
#   labels = scales::trans_format("log10", scales::math_format(10^.x))
# ) +
#theme(panel.grid.minor = element_blank())
#cowplot::plot_grid(degree_dens, eff_dens, sensitivity_dens, ncol=3,labels=c('A','B','C'))
#ggsave('tmp.png')

#    coord_trans(x = "log10", y = "log1p")
#    scale_x_continuous(trans = "log10")
    #scale_x_log10()
#    geom_histogram(binwidth=1)+
#    xlim(1,NA)+
#    ylim(NA, 0.1)
ggsave('tmp.png')

```

```{r}
real_random_labels = c('Real\nnetwork','Rewired\nnetwork')
names(real_random_labels)  = c('real','random')
pcc_deg_eff = pcc_df %>% filter(group=='real')%>%#pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
    ggplot(aes(x=deg, y=eff_norm))+
    geom_point()+
#    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
    theme_bw()+
    ylab("Effectiveness")+
    xlab("Degree")+
    stat_cor(method='pearson', cor.coef.name='R', label.x.npc=0.1) +
    theme(strip.text.y = element_text(angle = 0))+
    scale_y_continuous(labels = scales::label_scientific(digits=1))
pcc_deg_sens = pcc_df %>% filter(group=='real')%>%#pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
    ggplot(aes(x=deg, y=sens_norm))+
    geom_point()+
#    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
    theme_bw()+
    ylab("Sensitivity")+
    xlab("Degree")+
    stat_cor(method='spearman', cor.coef.name='rho', label.x.npc=0.1) +
    theme(strip.text.y = element_text(angle = 0))+
    scale_y_continuous(labels = scales::label_scientific(digits=1))
yuri_deg_eff = yuri_df %>% filter(group=='real')%>%#pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
    ggplot(aes(x=deg, y=eff_norm))+
    geom_point()+
#    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
    theme_bw()+
    ylab("Effectiveness")+
    xlab("Degree")+
    stat_cor(method='pearson', cor.coef.name='R', label.x.npc=0.1) +
    theme(strip.text.y = element_text(angle = 0))+
    scale_y_continuous(labels = scales::label_scientific(digits=1))
yuri_deg_sens = yuri_df %>% filter(group=='real')%>%#pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
    ggplot(aes(x=deg, y=sens_norm))+
    geom_point()+
#    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
    theme_bw()+
    ylab("Sensitivity")+
    xlab("Degree")+
    stat_cor(method='spearman', cor.coef.name='rho', label.x.npc=0.1) +
    theme(strip.text.y = element_text(angle = 0))+
    scale_y_continuous(labels = scales::label_scientific(digits=1))
huri_deg_eff = huri_df %>% filter(group=='real')%>%#pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
    ggplot(aes(x=deg, y=eff_norm))+
    geom_point()+
#    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
    theme_bw()+
    ylab("Effectiveness")+
    xlab("Degree")+
    stat_cor(method='pearson', cor.coef.name='R', label.x.npc=0.1) +
    theme(strip.text.y = element_text(angle = 0))+
    scale_y_continuous(labels = scales::label_scientific(digits=1))
huri_deg_sens = huri_df %>%filter(group=='real')%>% #pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
    ggplot(aes(x=deg, y=sens_norm))+
    geom_point()+
#    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) )+
    theme_bw()+
    ylab("Sensitivity")+
    xlab("Degree")+
    stat_cor(method='spearman', cor.coef.name='rho', label.x.npc=0.1) +
    theme(strip.text.y = element_text(angle = 0))+
    scale_y_continuous(labels = scales::label_scientific(digits=1))
deg_scatter_plots <- cowplot::plot_grid(pcc_deg_eff, pcc_deg_sens,
                   yuri_deg_eff, yuri_deg_sens,
                   huri_deg_eff, huri_deg_sens, labels = c('A','','B','','C',''),
                   ncol=2)
ggsave('tmp.png',plot=deg_scatter_plots, width=10,height=10)

#p1 = ggplot(pcc_df, aes(x=deg,y=eff_norm))+
#    geom_point(aes(color=cat_))+
#    xlab('Degree')+
#    ylab('Effectiveness')+
#    stat_cor(method='pearson')   +
#    guides(color=guide_legend(title="Experiment categories"))+
#    theme_bw()
#p2 = ggplot(pcc_df, aes(x=deg,y=sens_norm))+
#    geom_point(aes(color=cat_))+
#    xlab('Degree')+
#    ylab('Sensitivity')+
#    stat_cor(method='spearman')   +
#    guides(color=guide_legend(title="Experiment categories"))+
#    theme_bw()+theme(legend.text = element_text(margin = margin(t=5, b = 5, unit = "pt")))
#lgn = cowplot::get_legend(p2)
#fig3 = plot_grid(p1+theme(legend.position=''),
#          p2+theme(legend.position=''),
#          lgn, nrow=1, labels=c('A','B',''),rel_widths=c(0.4,0.4,0.2))
#ggsave('reports/figures/paper_figures_0916/fig3.png', fig3, width=8,height=6)
```

```{r}
pcc_rewire_df = read_csv('data/interim/pcc_0909/rewire_data.csv')
yuri_rewire_df = read_csv('data/interim/yuri/rewire_data.csv')
huri_rewire_df = read_csv('data/interim/huri/rewire_data.csv')

rewire_df = pcc_rewire_df%>%mutate(data='pcc')%>%bind_rows(
                                                           yuri_rewire_df%>%mutate(data='yuri'))%>%
        bind_rows(huri_rewire_df%>%mutate(data='huri'))
```

```{r}
network_labels  <- c('Costanzo 2016','Yeast PPI','HuRI')
#p3 = ggplot(rewire_df, aes(x=eff_deg_pearson))+
#    stat_density(geom = "line", trim=TRUE)+
##    geom_histogram(binwidth = 0.001)+
#    geom_vline(xintercept=0.9,color='red')+
#    theme_bw()

facet_labels  <- network_labels
names(facet_labels) <- c("pcc",'yuri','huri')

combined_rewire_df <- rewire_df%>%select(data, eff_deg_pearson,sens_deg_spearman)%>%mutate(type = 'random')%>% 
    bind_rows(
    data.frame(sens_deg_spearman= c(cor(pcc_df[pcc_df$group=='real',]$deg, pcc_df[pcc_df$group=='real',]$sens,method='spearman'),cor(yuri_df[yuri_df$group=='real',]$deg,yuri_df[yuri_df$group=='real',]$sens,method='spearman'),cor(huri_df[huri_df$group=='real',]$deg,huri_df[huri_df$group=='real',]$sens,method='spearman')), eff_deg_pearson= c(cor(pcc_df[pcc_df$group=='real',]$deg, pcc_df[pcc_df$group=='real',]$eff,method='pearson'),cor(yuri_df[yuri_df$group=='real',]$deg,yuri_df[yuri_df$group=='real',]$eff,method='pearson'),cor(huri_df[huri_df$group=='real',]$deg,huri_df[huri_df$group=='real',]$eff,method='pearson')), type='real',data=c('pcc','yuri','huri')))%>%
    mutate(data= fct_relevel(data, c('pcc','yuri','huri')))

p3 = 
    ggplot()+
    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
#    geom_boxplot()+
#    geom_jitter()+
    geom_density(data =combined_rewire_df%>%filter(type!='real'),aes(x=eff_deg_pearson,color=type, y=..scaled..),linetype = "dashed")+
    geom_vline(data =combined_rewire_df%>%filter(type=='real'),aes(xintercept=eff_deg_pearson,color=type))+
    guides(color=guide_legend(title="Data type"),shape=guide_legend(title="Network type"))+
    ylab('Density')+xlab('Effectiveness degree correlation')+
    theme_bw()+
    theme(panel.spacing = unit(2, "lines"))+
    scale_y_continuous(expand = c(0, 0)) 
p4 = 
    ggplot()+
    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
#    geom_boxplot()+
#    geom_jitter()+
    geom_density(data =combined_rewire_df%>%filter(type!='real'),aes(x=sens_deg_spearman,y=..scaled..,color=type),linetype = "dashed")+
    geom_vline(data =combined_rewire_df%>%filter(type=='real'),aes(xintercept=sens_deg_spearman,color=type))+
    guides(color=guide_legend(title="Data type"),shape=guide_legend(title="Network type"))+
    ylab('Density')+xlab('Sensitivity degree correlation')+
    theme_bw()+
    theme(panel.spacing = unit(2, "lines"),legend.position='bottom')+
    scale_color_discrete(labels=c('Rewired networks', 'Real network'))+
    scale_y_continuous(expand = c(0, 0)) 
lgn34  <- cowplot::get_legend(p4)
distr_plots <- plot_grid(
          p3+theme(legend.position=''),
 p4+theme(legend.position=''), lgn34,
 nrow=3, labels=c('D','E','','F','G',''), rel_heights = c(0.45,0.45,0.1))
#ggsave('reports/figures/paper_figures_0916/fig2.png',plot=distr_plots , width=7)

plot_grid(deg_scatter_plots, distr_plots, nrow=2,rel_heights = c(0.6,0.4))
ggsave('reports/figures/paper_figures_0916/fig2.png',width=8,height=9)
```

```{r deletion}
pcc_deletion = read_csv('data/interim/deletion_analysis_0911/pcc_deletion_100.csv')%>%
    pivot_wider(id_cols=c('range','variable_0'),names_from='variable_1',values_from='value')%>%
    mutate(range = range/max(range), gc_size = gc_size/max(gc_size), group='pcc')
yuri_deletion = read_csv('data/interim/deletion_analysis_0911/yuri_deletion_100.csv')%>%
    pivot_wider(id_cols=c('range','variable_0'),names_from='variable_1',values_from='value')%>%
    mutate(range = range/max(range), gc_size = gc_size/max(gc_size), group='yuri')
huri_deletion = read_csv('data/interim/deletion_analysis_0911/huri_deletion_100.csv')%>%
    pivot_wider(id_cols=c('range','variable_0'),names_from='variable_1',values_from='value')%>%
    mutate(range = range/max(range), gc_size = gc_size/max(gc_size), group='huri')

deletion_df  <- bind_rows(pcc_deletion, yuri_deletion, huri_deletion)%>%mutate(group = fct_relevel(group,c('pcc','yuri','huri')))

library(MESS)
auc_data <- deletion_df%>%
    group_by(group, variable_0)%>%
    summarise(auc=MESS::auc(gc_size,range), r = range[min(which(gc_size < 0.05))])%>%
    ungroup()%>%
    mutate(variable_0=fct_reorder(variable_0, auc, min))%>%pivot_wider(-r, names_from=group, values_from=auc)

auc_plot_pcc <- auc_data%>%
    ggplot(aes(x=variable_0,y=pcc))+
    geom_bar(aes(fill=variable_0,alpha = (variable_0 == "eff_sorted_deletion" )),stat='identity')+
    theme_bw()+
    ylab("Robustness")+
    xlab("Deletion parameter")+
    scale_alpha_manual(name = '', values = c(0.3, 1), guide = FALSE)+
    labs(fill = "Deletion\nparameter")+
    scale_x_discrete(labels = c(
                                "D",
                                "B",
                                "Eff",
                                "C",
                                "E",
                                "R",
                                "DR",
                                "S"))+
    scale_fill_brewer(name = 'Deletion\nparameter', palette = "Set1", guide='legend', labels= c(
                                "Degree",
                                "Betweenness",
                                "Effectiveness",
                                "Closeness",
                                "Eigencentrality",
                                "Random",
                                "Degree Reversed",
                                "Sensitivity"
                                ))+
   theme(legend.position='bottom', legend.title=element_text(size=18), legend.text=element_text(size=14),legend.spacing.x = unit(1.0, 'cm'),legend.spacing.y = unit(1.0, 'cm'))
auc_plot_yuri <- auc_data%>%
    ggplot(aes(x=variable_0,y=yuri))+
    geom_bar(aes(fill=variable_0,alpha = (variable_0 == "eff_sorted_deletion" | variable_0 == 'sens_sorted_deletion') ),stat='identity')+
    theme_bw()+
    ylab("Robustness")+
    xlab("Deletion parameter")+
    labs(fill = "Deletion\nparameter")+
    scale_alpha_manual(name = '', values = c(0.3, 1), guide = FALSE)+
    scale_x_discrete(labels = c(
                                "D",
                                "B",
                                "Eff",
                                "C",
                                "E",
                                "R", "DR",
                                "S"))+
    scale_fill_brewer(name='', palette = "Set1", labels= c(
                                "Degree",
                                "Betweenness",
                                "Effectiveness",
                                "Closeness",
                                "Eigencentrality",
                                "Random",
                                "Sensitivity","Degree Reversed"))+
   theme(legend.position='right')
auc_plot_huri<- auc_data%>%
    ggplot(aes(x=variable_0,y=huri, alpha = (variable_0 == "eff_sorted_deletion" ) ))+
    geom_bar(aes(fill=variable_0),stat='identity')+
    theme_bw()+
    scale_alpha_manual(values = c(0.3, 1), guide = FALSE)+
    ylab("Robustness")+
    xlab("Deletion parameter")+
    labs(fill = "Deletion\nparameter")+
    scale_x_discrete(labels = c(
                                "D",
                                "B",
                                "Eff",
                                "C",
                                "E",
                                "R",
                                "DR","S"))+
    scale_fill_brewer(palette = "Set1", labels= c(
                                "Degree",
                                "Betweenness",
                                "Effectiveness",
                                "Closeness",
                                "Eigencentrality",
                                "Random",
                                "Sensitivity","Degree Reversed"))+
   theme(legend.position='right')
#combined_df %>% group_by(data)%>%summarise(deg_btw = cor(deg, btw),
#                                           deg_eff = cor(deg,eff),
#                                           deg_sens = cor(deg,sens),
#                                           deg_eigen = cor(deg, eigenvec_centr),
#                                           deg_closeness = cor(deg, closeness_centr))
deletion_plot_pcc <-  pcc_deletion%>%   mutate(variable_0 = fct_relevel(variable_0, levels(auc_data$variable_0)))%>%
    ggplot(aes(x=range, y= gc_size, color = variable_0 , alpha = (variable_0 == "eff_sorted_deletion" ) ))+
#    geom_point(size=1)+
        geom_line(size=2)+
#    facet_grid(rows = vars(group), labeller = labeller(group=facet_labels))+
    scale_colour_brewer(palette = "Set1")+ 
    xlab('% Deleted Nodes')+
    ylab('% Giant Component Size')+
    scale_y_continuous(labels = scales::percent)+
    scale_x_continuous(labels = scales::percent)+
    theme_bw()+
    ggtitle('Costanzo 2016')+
    scale_alpha_manual(values = c(0.3, 1), guide = FALSE)+
    theme(legend.position='bottom',panel.border = element_blank(), axis.line = element_line())
deletion_plot_yuri <-  yuri_deletion%>%   mutate(variable_0 = fct_relevel(variable_0, levels(auc_data$variable_0)))%>%
ggplot(aes(x=range, y= gc_size, color = variable_0 ,alpha = (variable_0 == "eff_sorted_deletion" )  ))+
#    geom_point(size=1)+
        geom_line(size=2)+
#    facet_grid(rows = vars(group), labeller = labeller(group=facet_labels))+
    scale_colour_brewer(palette = "Set1")+ 
    xlab('% Deleted Nodes')+
    ggtitle('Yeast PPI')+
    ylab('% Giant Component Size')+
    scale_y_continuous(labels = scales::percent)+
    scale_x_continuous(labels = scales::percent)+
    scale_alpha_manual(values = c(0.3, 1), guide = FALSE)+
    theme_bw()+
    theme(legend.position='right',panel.border = element_blank(), axis.line = element_line())
deletion_plot_huri <-  huri_deletion%>%   mutate(variable_0 = fct_relevel(variable_0, levels(auc_data$variable_0)))%>%
ggplot(aes(x=range, y= gc_size, color = variable_0 , alpha = (variable_0 == "eff_sorted_deletion"  ) ))+
#    geom_point(size=1)+
        geom_line(size=2)+
#    facet_grid(rows = vars(group), labeller = labeller(group=facet_labels))+
    scale_colour_brewer(palette = "Set1")+ 
    xlab('% Deleted Nodes')+
    ylab('% Giant Component Size')+
    ggtitle('HuRI')+
    scale_y_continuous(labels = scales::percent)+
    scale_x_continuous(labels = scales::percent)+
    scale_alpha_manual(values = c(0.3, 1), guide = FALSE)+
    theme_bw()+
    theme(legend.position='right',panel.border = element_blank(), axis.line = element_line())
deletion_lgn <- cowplot::get_legend(auc_plot_pcc)
pcc_with_inset = ggdraw(deletion_plot_pcc+    theme(legend.position=''))+
  draw_plot(auc_plot_pcc+theme(legend.position=''), .50, .62, .45, .35) +
  draw_plot_label(
    c("A", ""),
    c(0, 0.45),
    c(1, 0.95),
    size = 12
  )
yuri_with_inset = ggdraw(deletion_plot_yuri+    theme(legend.position=''))+
  draw_plot(auc_plot_yuri+theme(legend.position=''), .50, .62, .45, .35) +
  draw_plot_label(
    c("B", ""),
    c(0, 0.45),
    c(1, 0.95),
    size = 12
  )
huri_with_inset = ggdraw(deletion_plot_huri+    theme(legend.position=''))+
  draw_plot(auc_plot_huri+theme(legend.position=''), .50, .62, .45, .35) +
  draw_plot_label(
    c("C", ""),
    c(0, 0.45),
    c(1, 0.95),
    size = 12
  )
plot_grid(pcc_with_inset, yuri_with_inset, huri_with_inset, ncol=3)%>%plot_grid(deletion_lgn,nrow=2, rel_heights=c(0.9,.1))
ggsave('reports/figures/paper_figures_0916/fig4.png',width=12)

plot_grid(deletion_plot_pcc+    theme(legend.position='')
, auc_plot_pcc+    theme(legend.position='')
,deletion_lgn, nrow=1, rel_widths =c(0.4, 0.5,.1), labels=c('A','B'))

```

# Sensor connectivity and neighbor degree

```{r}
sensor_connectivity_df <- read_csv('data/interim/sensor_connectivity_0915/sensor_connectivity_df.csv')%>%
    mutate(data= fct_relevel(data, c('pcc','yuri','huri')))

#library(ggridges)

#sensor_connectivity_df %>%
#    filter(type=='random')%>%
#    ggplot(aes(x=ratio,y=data))+
##    facet_grid(rows=vars(data))+
##    geom_histogram(bins=50)
#    geom_density_ridges()
#ggsave('tmp.png',width=12)

#p3 = 
sensor_connectivity_plot <- 
    ggplot()+
    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
#    geom_boxplot()+
#    geom_jitter()+
    geom_density(data =sensor_connectivity_df%>%filter(type!='real'),aes(x=ratio,color=type, y=..scaled..),linetype = "dashed")+
    geom_vline(data =sensor_connectivity_df%>%filter(type=='real'),aes(xintercept=ratio,color=type))+
    guides(color=guide_legend(title="Data type"),shape=guide_legend(title="Network type"))+
    ylab('Density')+xlab('Ratio of between group edges to outgroup edges')+
    theme_bw()+
    scale_color_discrete(labels=c('Random sampled set', 'Sensors'))+
    theme(legend.position='bottom')+
    scale_y_continuous(expand = c(0, 0))

#p3 = 
    ggplot()+
    geom_boxplot(data=sensor_connectivity_df%>%filter(type=='random'),
                 aes(x=data,y=ratio,color=type))+
#    geom_jitter()+
    geom_vline(data=sensor_connectivity_df%>%filter(type=='real'),
            aes(x=data,y=ratio,color=type, group=data))+
               #position=position_jitter(), aes(shape=data))+
    scale_color_discrete(labels=c('Random sampled set', 'Sensors'))+
    scale_x_discrete(labels=network_labels)+
  guides(color=guide_legend(title="Data type"),shape=guide_legend(title="Network type"))+
    xlab('Network type')+ylab('Ratio of between group edges\nto outgroup edges')+
    theme_bw()

neighbor_degree_plot <- combined_df%>%filter(group=='real')%>%group_by(data)%>%
    mutate(is_sensor = factor(ifelse(sens_norm>quantile(sens_norm,0.99),TRUE,FALSE), levels=c(TRUE,FALSE)))%>%
#    mutate(is_sensor = fct_relevel(is_sensor,c(TRUE,FALSE)))%>%
    ggplot(aes(x=is_sensor,y=log(neighbor_degree)))+
    facet_grid(cols=vars(data),labeller = labeller(data=facet_labels))+
#    geom_point()+
    geom_boxplot()+
    theme_bw()+
    ylab('Log(Average neighbor degree)')+
    xlab('')+ stat_compare_means()+
    scale_x_discrete(labels=c('aensor', 'Not Sensor'))
plot_grid(neighbor_degree_plot, sensor_connectivity_plot, nrow=2 , labels=c('A','B'))
ggsave('reports/figures/paper_figures_0916/fig5.png')

#combined_df%>%filter(group=='real'&data=='huri')%>%mutate(is_sensor = ifelse(sens_norm>quantile(sens_norm,0.99),TRUE,FALSE))%>%
#    coin::independence_test(neighbor_degree~is_sensor, data=.)

```

```{r}
convert_networkx_to_igraph <- 			# 
    function
(
 	nx_object,
    df
)
{
    nx =  reticulate::import("networkx")
    ig = igraph::graph.adjacency(nx$adjacency_matrix(nx_object), mode = 'undirected')
    #ig = igraph::set_vertex_attr(ig , name =  'orf_name', value  = (df$orf_name))
    ig = igraph::set_vertex_attr(ig , name =  'name', value  = (df$orf_name))
    #igraph::set_vertex_attr(ig, "pos", nx_object$nodes('pos'))
    ig
}
```