### 2020-12-21 notebooks/10-enm_rnaseq-0825.Rmd
notebooks with all analyses i have until Dec 21st 2020

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.5.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

<!-- #region toc=true -->
<h1>Table of Contents<span class="tocSkip"></span></h1>
<div class="toc"><ul class="toc-item"></ul></div>
<!-- #endregion -->

```{r}
library(tidyverse)
library(edgeR)
library(tximport)
library(GenomicFeatures)
```

```{r}
ann  <-  '~/data/genome/gtf/Saccharomyces_cerevisiae.R64-1-1.46.gtf'
table_  <- data.table::fread('~/data/PRJNA475469/SraRunTable_PRJNA475469.txt')
samples <- table_ %>%
                dplyr::select(`Run`,`kinase_inhibitor`,
                       `media`, stress, Genotype)%>%
                #unite('condition', `Factor Value[genotype]`,`Factor Value[environmental stress]`)%>%
                distinct()#
```

```{r}
files <- file.path('~/data/rnaseq/PRJNA475469/quants',(samples$Run),'quant.sf')
names(files) <- samples$Run
```

```{r jupyter={'outputs_hidden': True}}
txdb <- makeTxDbFromEnsembl("Saccharomyces cerevisiae",
                                    server="useastdb.ensembl.org")
k <- keys(txdb, keytype = "TXNAME")
tx2gene <- AnnotationDbi::select(txdb, k, "GENEID", "TXNAME")
txi.salmon <- tximport(files,type='salmon',tx2gene=tx2gene)

ypd_wt <-  samples%>%filter(Genotype=='WT')%>%mutate(stress=ifelse(media=='SDC','SD',ifelse(stress=='None','YPD',stress)))
ypd_wt_runids <-ypd_wt%>%dplyr::select(Run)%>%pull()
```

```{r}
counts_mat<- txi.salmon$counts[,ypd_wt_runids]
y <- DGEList(counts=counts_mat)
y <- calcNormFactors(y,method='TMM')
tmm <- cpm(y)
```

```{r}
samples
```

```{r}

```

```{r}
ypd_sdc_runids <- samples%>%filter(kinase_inhibitor=='None'&(stress=='None'|stress=='Glucose_Dropout')&media=='YPD'&Genotype=='WT')%>%dplyr::select(Run)%>%pull()
ypd_sdc_runids
```

```{r}

counts_mat<- txi.salmon$counts[,ypd_sdc_runids]
y <- DGEList(counts=counts_mat)
y <- calcNormFactors(y,method='TMM')
tmm <- cpm(y)

condition <- factor(samples%>%dplyr::filter(Run%in%ypd_sdc_runids)%>%dplyr::select(stress)%>%pull())
condition<-relevel(condition,ref='None')
design <- model.matrix(~condition)
rownames(design) <- colnames(y)
y <- estimateDisp(y,design,robust=T)
y$common.dispersion

```

```{r}
fit <- glmFit(y,design)
lrt_ds <- glmLRT(fit,coef=2)
FDR <- p.adjust(lrt_ds$table$PValue, method="BH")
tb <- lrt_ds$table
tb['FDR']  <- FDR

#tb[lst_of_genes,]
```

```{r}
enm_df <- data.table::fread('../data/interim/0824_df_withlinkcom.csv')
```

```{r jupyter={'outputs_hidden': True}}
enm_df
```

```{r}
strain_ids <- data.table::fread('../data/interim/strain_ids_with_experiment_count_all.csv')
```

```{r}
enm_rnaseq = enm_df %>% 
inner_join(strain_ids,c('orf_name'='Allele Gene name'))%>%
inner_join(tb%>%rownames_to_column('orf_name'),c('Systematic gene name.y'='orf_name'))
```

```{r}
enm_rnaseq %>%ggplot(aes(x=eff,y=(logFC),col=FDR < 0.01))+
    geom_point()+
    geom_hline(yintercept=1)+
    geom_hline(yintercept=-1)
```

```{r}

```