### 2021-04-05 notebooks/paper_plots_1213_pcc_only.Rmd
text size changes for fig5

---
title: "Paper figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
# use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv("enm")
setwd("../")
source("src/R/plot_density_prs.R")
figure_path = 'reports/figures/0404_paperfigure_sensors/'
```

```{python}
import matplotlib.pyplot as plt
import matplotlib as mpl
import networkx as nx
import numpy as np
import pandas as pd
import plotnine as p9
import pickle 
import os
import re
import itertools as itr
from src.enm import Enm

with open(f'data/interim/pcc_0909/pcc.pickle','rb') as f:
    e_pcc = pickle.load(f)
pcc_df = e_pcc.df
pcc_gc = e_pcc.graph_gc
pcc_gc_pos = (dict(pcc_gc.nodes(data='pos')))
```

```{r}
library(tidyverse)
library(glue)
library(cowplot)
library(ggpubr)
library(ggsignif)

pcc_df_random <- data.table::rbindlist(lapply(py$e_pcc$e_list, function(x) x$df), fill = T, idcol = "random_id") %>% mutate(random_id = as.factor(random_id))
pcc_df <- py$pcc_df %>%
  mutate(data = "pcc") %>%
  mutate(sens_norm = sens / nrow(.), eff_norm = eff / nrow(.), group = "real") %>%
  bind_rows(
    pcc_df_random %>% mutate(data = "pcc") %>% mutate(sens_norm = sens / nrow(py$pcc_df), eff_norm = eff / nrow(py$pcc_df), group = "random")
  ) %>%
  mutate(group = fct_relevel(group, c("real", "random")))
combined_df <- pcc_df
```

Degree - Eff - Sens distributions
```{r}

pcc_eff_dens <-
  plot_density_prs(pcc_df, eff_norm, xlab = "Effectiveness")
dens_lgn <- cowplot::get_legend(pcc_eff_dens + theme(legend.position = "right"))
pcc_degree_dens <- cowplot::ggdraw(plot_density_prs(pcc_df, deg, xlab = "Degree") + theme(legend.position = "none")) #+
#  cowplot::draw_plot(dens_lgn, .55,.55,.5,.5)
pcc_eff_dens <-
  cowplot::ggdraw(plot_density_prs(pcc_df, eff_norm, xlab = "Effectiveness") + theme(legend.position = "none") +
    coord_trans(x = "log10", xlim = c(NA, 10e-2)) +
    scale_x_continuous(
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    )) #+
#   cowplot::draw_plot(dens_lgn, .55,.55,.5,.5)
# annotate("text", x = 0.000001, y = 0.2, label=sprintf("p = %#.2f", eff_p), fontface='italic', size = 6)
pcc_sensitivity_dens <-
  cowplot::ggdraw(plot_density_prs(pcc_df, sens_norm, xlab = "Sensitivity") + theme(legend.position = "none")) #+
#        cowplot::draw_plot(dens_lgn, .55,.55,.5,.5)
# title <- cowplot::ggdraw() + cowplot::draw_label("Yeast GI Profile Similarity", fontface='bold')
pcc_dens_comb <- cowplot::plot_grid(pcc_eff_dens + theme(legend.position = "none"),
  pcc_sensitivity_dens + theme(legend.position = "none"),
  #+theme(legend.position='none'),
  labels = c("F", "G"), ncol = 2
)
# pcc_dens_comb_title = cowplot::plot_grid(title, pcc_dens_comb, nrow=2 ,rel_heights=c(0.1,0.9))
# dens_lgn = cowplot::get_legend(pcc_sensitivity_dens)
dens_plots <- pcc_dens_comb # cowplot::plot_grid(pcc_dens_comb_title,nrow=1)
# cowplot::plot_grid(dens_plots, dens_lgn,nrow=2,rel_heights=c(.9,.1))
# ggsave('reports/figures/paper_figures_0202_pcc_only/fig1.png', plot=pcc_dens_comb,width=4,height=4)
```

```{r}
real_random_labels <- c("PSN", "Rewired\nnetwork")
names(real_random_labels) <- c("real", "random")
pcc_deg_eff <- pcc_df %>% filter(group == "real") %>% # pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
  ggplot(aes(x = deg, y = eff_norm)) +
  geom_point() +
  #    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
  theme_bw() +
  ylab("Effectiveness") +
  xlab("Degree") +
  stat_cor(method = "pearson", cor.coef.name = "R", label.x.npc = 0.1) +
  theme(strip.text.y = element_text(angle = 0)) +
  scale_y_continuous(labels = scales::label_scientific(digits = 1))
pcc_deg_sens <- pcc_df %>% filter(group == "real") %>% # pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
  ggplot(aes(x = deg, y = sens_norm)) +
  geom_point() +
  #    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
  theme_bw() +
  ylab("Sensitivity") +
  xlab("Degree") +
  stat_cor(method = "spearman", cor.coef.name = "rho", label.x.npc = 0.1) +
  theme(strip.text.y = element_text(angle = 0)) +
  scale_y_continuous(labels = scales::label_scientific(digits = 1))
deg_scatter_plots <- cowplot::plot_grid(pcc_deg_eff, pcc_deg_sens,
  labels = c("A", "B"),
  ncol = 2
)
ggsave("reports/figures/paper_figures_0202_pcc_only/fig2.png", plot = deg_scatter_plots, width = 10, height = 10)
```

```{r}
pcc_rewire_df <- read_csv("data/interim/pcc_0909/rewire_data.csv")
rewire_df <- pcc_rewire_df %>% mutate(data = "pcc")
network_labels <- c("Yeast GI Profile Similarity")
facet_labels <- network_labels
names(facet_labels) <- c("pcc")
combined_rewire_df <- rewire_df %>%
  select(data, eff_deg_pearson, sens_deg_spearman) %>%
  mutate(type = "random") %>%
  bind_rows(
    data.frame(
      sens_deg_spearman = c(cor(pcc_df[pcc_df$group == "real", ]$deg, pcc_df[pcc_df$group == "real", ]$sens, method = "spearman")),
      eff_deg_pearson = c(cor(pcc_df[pcc_df$group == "real", ]$deg, pcc_df[pcc_df$group == "real", ]$eff, method = "pearson")), type = "real", data = c("pcc")
    )
  ) %>%
  mutate(data = fct_relevel(data, c("pcc")))

p3 <-
  ggplot() +
  #    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
  #    geom_boxplot()+
  #    geom_jitter()+
  geom_density(data = combined_rewire_df %>% filter(type != "real"), aes(x = eff_deg_pearson, color = type, y = ..scaled..), linetype = "dashed") +
  geom_vline(data = combined_rewire_df %>% filter(type == "real"), aes(xintercept = eff_deg_pearson, color = type)) +
  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
  ylab("Density") +
  xlab("Effectiveness degree correlation") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines")) +
  scale_y_continuous(expand = c(0, 0))
p4 <-
  ggplot() +
  #    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
  #    geom_boxplot()+
  #    geom_jitter()+
  geom_density(data = combined_rewire_df %>% filter(type != "real"), aes(x = sens_deg_spearman, y = ..scaled.., color = type), linetype = "dashed") +
  geom_vline(data = combined_rewire_df %>% filter(type == "real"), aes(xintercept = sens_deg_spearman, color = type)) +
  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
  ylab("Density") +
  xlab("Sensitivity degree correlation") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines"), legend.position = "bottom") +
  scale_color_discrete(labels = c("Random networks", "Genetic network")) +
  scale_y_continuous(expand = c(0, 0))
lgn34 <- cowplot::get_legend(p4)
distr_plots <- plot_grid(pcc_degree_dens + theme(legend.position = ""),
  ggdraw(p3 + theme(legend.position = "")),
  ggdraw(p4 + theme(legend.position = "")),
  labels = c("C", "D", "E"), nrow = 1
)
# distr_plots_withlgn <- plot_grid(distr_plots, lgn34,
# nrow=2,rel_heights = c(0.9,0.1))
# ggsave('reports/figures/paper_figures_0916/fig2.png',plot=distr_plots , width=7)
plot_grid(deg_scatter_plots, distr_plots, dens_plots, nrow = 3, rel_heights = c(0.3, 0.4, .3))
ggsave("reports/figures/paper_figures_0202_pcc_only/fig3.pdf", width = 10, height = 12)
```

```{r deletion}
pcc_deletion <- read_csv("data/interim/deletion_analysis_0911/pcc_deletion_100.csv") %>%
  pivot_wider(id_cols = c("range", "variable_0"), names_from = "variable_1", values_from = "value") %>%
  mutate(range = range / max(range), gc_size = gc_size / max(gc_size), group = "pcc")
yuri_deletion <- read_csv("data/interim/deletion_analysis_0911/yuri_deletion_100.csv") %>%
  pivot_wider(id_cols = c("range", "variable_0"), names_from = "variable_1", values_from = "value") %>%
  mutate(range = range / max(range), gc_size = gc_size / max(gc_size), group = "yuri")
huri_deletion <- read_csv("data/interim/deletion_analysis_0911/huri_deletion_100.csv") %>%
  pivot_wider(id_cols = c("range", "variable_0"), names_from = "variable_1", values_from = "value") %>%
  mutate(range = range / max(range), gc_size = gc_size / max(gc_size), group = "huri")

deletion_df <- bind_rows(pcc_deletion, yuri_deletion, huri_deletion) %>% mutate(group = fct_relevel(group, c("pcc", "yuri", "huri")))

library(MESS)
auc_data <- deletion_df %>%
  group_by(group, variable_0) %>%
  summarise(auc = MESS::auc(gc_size, range), r = range[min(which(gc_size < 0.05))]) %>%
  ungroup() %>%
  mutate(variable_0 = fct_reorder(variable_0, auc, min)) %>%
  pivot_wider(-r, names_from = group, values_from = auc)

auc_plot_pcc <- auc_data %>%
  ggplot(aes(x = variable_0, y = pcc)) +
  geom_bar(aes(fill = variable_0, alpha = (variable_0 == "eff_sorted_deletion")), stat = "identity") +
  theme_bw() +
  ylab("Robustness") +
  xlab("Deletion parameter") +
  scale_alpha_manual(name = "", values = c(0.3, 1), guide = FALSE) +
  labs(fill = "Deletion\nparameter") +
  scale_x_discrete(labels = c(
    "D",
    "B",
    "Eff",
    "C",
    "E",
    "R",
    "DR",
    "S"
  )) +
  scale_fill_brewer(name = "Deletion\nparameter", palette = "Set1", guide = "legend", labels = c(
    "Degree",
    "Betweenness",
    "Effectiveness",
    "Closeness",
    "Eigencentrality",
    "Random",
    "Degree Reversed",
    "Sensitivity"
  )) +
  theme(legend.position = "bottom", legend.title = element_text(size = 18), legend.text = element_text(size = 14), legend.spacing.x = unit(1.0, "cm"), legend.spacing.y = unit(1.0, "cm"))
auc_plot_yuri <- auc_data %>%
  ggplot(aes(x = variable_0, y = yuri)) +
  geom_bar(aes(fill = variable_0, alpha = (variable_0 == "eff_sorted_deletion" | variable_0 == "sens_sorted_deletion")), stat = "identity") +
  theme_bw() +
  ylab("Robustness") +
  xlab("Deletion parameter") +
  labs(fill = "Deletion\nparameter") +
  scale_alpha_manual(name = "", values = c(0.3, 1), guide = FALSE) +
  scale_x_discrete(labels = c(
    "D",
    "B",
    "Eff",
    "C",
    "E",
    "R", "DR",
    "S"
  )) +
  scale_fill_brewer(name = "", palette = "Set1", labels = c(
    "Degree",
    "Betweenness",
    "Effectiveness",
    "Closeness",
    "Eigencentrality",
    "Random",
    "Sensitivity", "Degree Reversed"
  )) +
  theme(legend.position = "right")
auc_plot_huri <- auc_data %>%
  ggplot(aes(x = variable_0, y = huri, alpha = (variable_0 == "eff_sorted_deletion"))) +
  geom_bar(aes(fill = variable_0), stat = "identity") +
  theme_bw() +
  scale_alpha_manual(values = c(0.3, 1), guide = FALSE) +
  ylab("Robustness") +
  xlab("Deletion parameter") +
  labs(fill = "Deletion\nparameter") +
  scale_x_discrete(labels = c(
    "D",
    "B",
    "Eff",
    "C",
    "E",
    "R",
    "DR", "S"
  )) +
  scale_fill_brewer(palette = "Set1", labels = c(
    "Degree",
    "Betweenness",
    "Effectiveness",
    "Closeness",
    "Eigencentrality",
    "Random",
    "Sensitivity", "Degree Reversed"
  )) +
  theme(legend.position = "right")
# combined_df %>% group_by(data)%>%summarise(deg_btw = cor(deg, btw),
#                                           deg_eff = cor(deg,eff),
#                                           deg_sens = cor(deg,sens),
#                                           deg_eigen = cor(deg, eigenvec_centr),
#                                           deg_closeness = cor(deg, closeness_centr))
deletion_plot_pcc <- pcc_deletion %>%
  mutate(variable_0 = fct_relevel(variable_0, levels(auc_data$variable_0))) %>%
  ggplot(aes(x = range, y = gc_size, color = variable_0, alpha = (variable_0 == "eff_sorted_deletion"))) +
  #    geom_point(size=1)+
  geom_line(size = 2) +
  #    facet_grid(rows = vars(group), labeller = labeller(group=facet_labels))+
  scale_colour_brewer(palette = "Set1") +
  xlab("% Deleted Nodes") +
  ylab("% Giant Component Size") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  ggtitle("Costanzo 2016") +
  scale_alpha_manual(values = c(0.3, 1), guide = FALSE) +
  theme(legend.position = "bottom", panel.border = element_blank(), axis.line = element_line())
deletion_plot_yuri <- yuri_deletion %>%
  mutate(variable_0 = fct_relevel(variable_0, levels(auc_data$variable_0))) %>%
  ggplot(aes(x = range, y = gc_size, color = variable_0, alpha = (variable_0 == "eff_sorted_deletion"))) +
  #    geom_point(size=1)+
  geom_line(size = 2) +
  #    facet_grid(rows = vars(group), labeller = labeller(group=facet_labels))+
  scale_colour_brewer(palette = "Set1") +
  xlab("% Deleted Nodes") +
  ggtitle("Yeast PPI") +
  ylab("% Giant Component Size") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  scale_alpha_manual(values = c(0.3, 1), guide = FALSE) +
  theme_bw() +
  theme(legend.position = "right", panel.border = element_blank(), axis.line = element_line())
deletion_plot_huri <- huri_deletion %>%
  mutate(variable_0 = fct_relevel(variable_0, levels(auc_data$variable_0))) %>%
  ggplot(aes(x = range, y = gc_size, color = variable_0, alpha = (variable_0 == "eff_sorted_deletion"))) +
  #    geom_point(size=1)+
  geom_line(size = 2) +
  #    facet_grid(rows = vars(group), labeller = labeller(group=facet_labels))+
  scale_colour_brewer(palette = "Set1") +
  xlab("% Deleted Nodes") +
  ylab("% Giant Component Size") +
  ggtitle("HuRI") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  scale_alpha_manual(values = c(0.3, 1), guide = FALSE) +
  theme_bw() +
  theme(legend.position = "right", panel.border = element_blank(), axis.line = element_line())
deletion_lgn <- cowplot::get_legend(auc_plot_pcc)
pcc_with_inset <- ggdraw(deletion_plot_pcc + theme(legend.position = "")) +
  draw_plot(auc_plot_pcc + theme(legend.position = ""), .50, .62, .45, .35) +
  draw_plot_label(
    c("A", ""),
    c(0, 0.45),
    c(1, 0.95),
    size = 12
  )
yuri_with_inset <- ggdraw(deletion_plot_yuri + theme(legend.position = "")) +
  draw_plot(auc_plot_yuri + theme(legend.position = ""), .50, .62, .45, .35) +
  draw_plot_label(
    c("B", ""),
    c(0, 0.45),
    c(1, 0.95),
    size = 12
  )
huri_with_inset <- ggdraw(deletion_plot_huri + theme(legend.position = "")) +
  draw_plot(auc_plot_huri + theme(legend.position = ""), .50, .62, .45, .35) +
  draw_plot_label(
    c("C", ""),
    c(0, 0.45),
    c(1, 0.95),
    size = 12
  )
plot_grid(pcc_with_inset, yuri_with_inset, huri_with_inset, ncol = 3) %>% plot_grid(deletion_lgn, nrow = 2, rel_heights = c(0.9, .1))
ggsave("reports/figures/paper_figures_0916/fig4.png", width = 12)

plot_grid(deletion_plot_pcc + theme(legend.position = ""),
  auc_plot_pcc + theme(legend.position = ""),
  deletion_lgn,
  nrow = 1, rel_widths = c(0.4, 0.5, .1), labels = c("A", "B")
)
```

# Sensor connectivity and neighbor degree

```{r}
sensor_connectivity_df <- read_csv("data/interim/sensor_connectivity_1205/sensor_connectivity_df.csv") %>%
  mutate(data = fct_relevel(data, c("pcc", "gi", "yuri", "huri"))) %>%
  filter(data == "pcc")
# effector_connectivity_df <- read_csv('data/interim/sensor_connectivity_1205/effector_connectivity_df.csv')%>%
#    mutate(data= fct_relevel(data, c('pcc','gi','yuri','huri')))%>%filter(data=='pcc')
# effector_connectivity_plot <-
#    ggplot()+
##    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
##    geom_boxplot()+
##    geom_jitter()+
#    geom_density(data =effector_connectivity_df%>%filter(type!='real'),aes(x=ratio,color=type, y=..scaled..),linetype = "dashed")+
#    geom_vline(data =effector_connectivity_df%>%filter(type=='real'),aes(xintercept=ratio,color=type))+
#    guides(color=guide_legend(title="Data type"),shape=guide_legend(title="Network type"))+
#    ylab('Density')+xlab('Ratio of between group edges to outgroup edges')+
#    theme_bw()+
#    scale_color_discrete(labels=c('Random sampled set', 'Effectors'))+
#    theme(legend.position='bottom')+
#    scale_y_continuous(expand = c(0, 0))
eff_neighbor_degree_plot <- combined_df %>%
  filter(group == "real") %>%
  group_by(data) %>%
  mutate(is_effector = factor(ifelse(eff_norm > quantile(eff_norm, 0.99), TRUE, FALSE), levels = c(TRUE, FALSE))) %>%
  #    mutate(is_effector = fct_relevel(is_effector,c(TRUE,FALSE)))%>%
  ggplot(aes(x = is_effector, y = log(neighbor_degree))) +
  #    facet_grid(cols=vars(data),labeller = labeller(data=facet_labels))+
  #    geom_point()+
  geom_boxplot() +
  theme_bw() +
  ylab("Log(Average neighbor degree)") +
  xlab("") +
  stat_compare_means() +
  scale_x_discrete(labels = c("Effector", "Not Effector"))
sensor_connectivity_plot <-
  ggplot() +
  #    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
  #    geom_boxplot()+
  #    geom_jitter()+
  geom_density(data = sensor_connectivity_df %>% filter(type != "real"), aes(x = ratio, color = type, y = ..scaled..), linetype = "dashed") +
  geom_vline(data = sensor_connectivity_df %>% filter(type == "real"), aes(xintercept = ratio, color = type)) +
  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
  ylab("Density") +
  xlab("Ratio of between group edges\nto outgroup edges") +
  theme_bw() +
  scale_color_discrete(labels = c("Random sampled set", "Sensors")) +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0, 0))
neighbor_degree_plot <- combined_df %>%
  filter(group == "real") %>%
  group_by(data) %>%
  mutate(is_sensor = factor(ifelse(sens_norm > quantile(sens_norm, 0.99), TRUE, FALSE), levels = c(TRUE, FALSE))) %>%
  #    mutate(is_sensor = fct_relevel(is_sensor,c(TRUE,FALSE)))%>%
  ggplot(aes(x = is_sensor, y = log(neighbor_degree))) +
  #    facet_grid(cols=vars(data),labeller = labeller(data=facet_labels))+
  #    geom_point()+
  geom_boxplot() +
  theme_bw() +
  ylab("Log(Average neighbor degree)") +
  xlab("") +
  stat_compare_means() +
  scale_x_discrete(labels = c("Sensor", "Not Sensor"))
sensor_conn_lgn <- get_legend(sensor_connectivity_plot + theme(legend.position = "right"))
plot_grid(neighbor_degree_plot,
  sensor_connectivity_plot + theme(legend.position = "none"), # effector_connectivity_plot,
  eff_neighbor_degree_plot, sensor_conn_lgn,
  nrow = 2,
  labels = c("A", "B", "C", "")
) # %>%
#    plot_grid(, labels=c('','C'),nrow=2, rel_widths=c(2,1))
ggsave("reports/figures/paper_figures_0202_pcc_only/fig5.pdf", width = 6, height = 6)

# combined_df%>%filter(group=='real'&data=='huri')%>%mutate(is_sensor = ifelse(sens_norm>quantile(sens_norm,0.99),TRUE,FALSE))%>%
#    coin::independence_test(neighbor_degree~is_sensor, data=.)
```

```{r}
convert_networkx_to_igraph <- #
  function
  (
   nx_object,
   df) {
    nx <- reticulate::import("networkx")
    ig <- igraph::graph.adjacency(nx$adjacency_matrix(nx_object), mode = "undirected")
    # ig = igraph::set_vertex_attr(ig , name =  'orf_name', value  = (df$orf_name))
    ig <- igraph::set_vertex_attr(ig, name = "name", value = (df$orf_name))
    # igraph::set_vertex_attr(ig, "pos", nx_object$nodes('pos'))
    ig
  }
```
```{r}
library(RColorBrewer)
col <- brewer.pal(n = 9, name = "Set1")
sensor_colors <- col[1:6] # brewer.pal(n = 6, name = "Set1")
effector_colors <- col[7:9] # brewer.pal(n=3, name= "Set1_r")
```

```{python}
sensor_colors = r.sensor_colors
effector_colors = r.effector_colors
sensors_pcc = e_pcc.df.loc[e_pcc.df.sens>np.quantile(e_pcc.df.sens,0.99)]
#sensor_sub_pcc = get_subnetwork(e_pcc.graph_gc, sensors_pcc.orf_name.values)
#is_sensor = dict(zip(sensor_sub_pcc.nodes, [True if i in sensors_pcc.orf_name.values else False  for i in sensor_sub_pcc.nodes]))
#nx.set_node_attributes(sensor_sub_pcc, is_sensor, 'is_sensor')
sensor_go_terms_separate = [
    'iron ion\ntransport',
    'phenylalanine\ntransport',
    None, 
    None,
    None,
    None,
    'tricarboxylic\nacid cycle', 
    'hexose metabolic\nprocess',
    'protein folding',
    'mitochondria-nucleus\nsignaling pathway/TCA',
    None,
    None,
    None,
    None,
    None,
    None,
    None,
]
sensor_connected_components = sorted([sorted(list(i)) for i in nx.connected_components(nx.induced_subgraph(e_pcc.graph_gc, sensors_pcc.orf_name))], key=lambda x: x[0])
sensor_go_terms= {}
for i in range(len(sensor_go_terms_separate)):
    for j in sensor_connected_components[i]:
        sensor_go_terms[j] = sensor_go_terms_separate[i]
sensors_pcc = sensors_pcc.merge(pd.DataFrame.from_dict(sensor_go_terms,orient='index',columns=['go_group']),right_index=True,left_on='orf_name')

effector_pcc = e_pcc.df.loc[e_pcc.df.eff>np.quantile(e_pcc.df.eff,0.99)]
#effector_sub_pcc = get_subnetwork(e_pcc.graph_gc, effector_pcc.orf_name.values)
#is_effector = dict(zip(effector_sub_pcc.nodes, [True if i in effector_pcc.orf_name.values else False  for i in effector_sub_pcc.nodes]))
#nx.set_node_attributes(effector_sub_pcc, is_effector, 'is_effector')
effectors_connected_comp = sorted([sorted(list(i)) for i in nx.connected_components(nx.induced_subgraph(e_pcc.graph_gc, effector_pcc.orf_name.tolist()))], key=lambda x:x[0])
effector_pcc.loc[:, 'go_group'] = ['mito' if i in effectors_connected_comp[0] else 'golgi' if i in effectors_connected_comp[1] else 'chromatin' for i in effector_pcc.orf_name]
effector_order = effector_pcc.groupby('go_group').eff.median().sort_values().index.tolist()
sensor_order = sensors_pcc.groupby('go_group').sens.median().sort_values().index.tolist()
```

# Network plot only sensors
```{python}
from matplotlib.lines import Line2D
pos = e_pcc.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#                         node_shape=['^' if i in sensors_pcc.orf_name.values else 'v' if i in effector_pcc.orf_name.values else 'o' for i in e_pcc.nodes])
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=sensors_pcc.orf_name.values, node_size=200, pos=pos,
                          node_color='black',
                          node_shape='^',edgecolors='black',
                          linewidths=1)
legend_elements.append(
    Line2D([0], [0], marker='^', color='black', label=f'Sensors (No GO enrichment)',
                          markerfacecolor=sensor_colors[0], markersize=12, linestyle="None")
)
for itr, i in enumerate(sensor_order):
    #print(i, effector_colors[itr])

    orf_names_to_plot = sensors_pcc.loc[sensors_pcc.go_group==i, 'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=sensor_colors[itr],
                          node_shape='^',edgecolors='black',
                          linewidths=1)
    legend_elements.append(
        Line2D([0], [0], marker='^', color='black', label=f'Sensors ({i})',
                              markerfacecolor=sensor_colors[itr], markersize=16, linestyle="None")
    )
nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, sensors_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='red', alpha=0.5)
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-"),
                   Line2D([0], [0], marker='o', color='red', label= 'Sensor-Sensor edges',
                              markerfacecolor='#018571', markersize=0, linestyle="-")
])
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))
figure_path = r.figure_path
plt.savefig(f'{figure_path}//tmp_nwsens.png',bbox_inches='tight')
#plt.savefig('tmp_nw.png',bbox_inches='tight')
```

```{python}
from matplotlib.lines import Line2D
pos = e_pcc.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#        Line2D([0], [0], marker='^', color='black', label=f'Sensors ({i})',
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=effector_pcc.orf_name.values, node_size=100, pos=pos,
                      node_color='black',
                      node_shape='o')
for itr, i in enumerate(effector_order):
    orf_names_to_plot = effector_pcc.loc[effector_pcc.go_group==i,'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=400, pos=pos,
                          node_color=effector_colors[itr],
                          node_shape='o',edgecolors='black',
                          linewidths=1)
    if i == 'mito':
        lbl = 'Respiratory complex assembly'
    elif i == 'golgi':
        lbl = "Golgi vesicle transport"
    else:
        lbl = 'Chromosome segregation'
    legend_elements.append(
            Line2D([0], [0], marker='o', color='black', label=f'Effectors ({lbl})',
                                  markerfacecolor=effector_colors[itr], markersize=16, linestyle="None")
        )
nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, effector_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='blue',alpha=0.5)
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-"),
#                   Line2D([0], [0], marker='o', color='red', label= 'Sensor-Sensor edges',
#                              markerfacecolor='#018571', markersize=0, linestyle="-")
                   Line2D([0], [0], marker='o', color='blue', label= 'Effector-Effector edges',
                              markerfacecolor='#a6611a', markersize=0, linestyle="-")
])
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))

nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, effector_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='blue',alpha=0.5)
plt.savefig('tmp_nw.png',bbox_inches='tight')
```

```{r}
effector_pcc <- py$effector_pcc
sensors_pcc <- py$sensors_pcc %>% mutate(go_group = (as.character(go_group)), go_group = ifelse(go_group == "NULL", NA, go_group))
# library(tidyverse)
effector_pcc$go_group <- as.factor(effector_pcc$go_group)

# res.aov <- kruskal.test(eff~go_group, data= effector_pcc)
# res.aov
effector_eff_pval <- FSA::dunnTest(eff ~ go_group, effector_pcc)
effector_eff_pval$res[effector_eff_pval$res$P.adj < 0.1, ]
effector_deg_pval <- FSA::dunnTest(deg ~ go_group, effector_pcc)
effector_deg_pval$res[effector_deg_pval$res$P.adj < 0.1, ]

eff_deg_comp <- ggplot(effector_pcc %>% right_join(py$e_pcc$df)%>% mutate(go_group = factor(replace_na(as.character(go_group),"Others")),go_group=fct_reorder(go_group, eff),go_group=fct_relevel(go_group,'Others',after=Inf)), 
                       aes(x = go_group, y = deg, fill = go_group)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(
    values = c(effector_colors,'black'),
    labels = c("Chromosome\nSegregetion", "Golgi vesicle\nTransport", "Respiratory complex\nassembly",'Non-effector genes')
  ) +
  xlab("") +
  ylab("Degree") +
  theme_bw() +
  geom_signif(
    comparisons = list(c("golgi", "chromatin"), c("golgi", "mito"), c("mito", "chromatin")),
    annotation = c("**", "***", "ns"),
    y_position = c(115, 120, 130)
  ) +
  theme(
       axis.text.y=element_text(size=10),
        axis.title.y=element_text(size=14), axis.text.x = element_blank(), legend.title = element_blank(), legend.text = element_text(margin = margin(t = 10, b = 10))) # , legend.spacing.y = unit(2,'in'))+
ggsave(glue('{figure_path}//tmp_eff_deg_nolegend.png'), plot = eff_deg_comp+theme(legend.position='none'),width=4,height=4)
ggsave(glue('{figure_path}//tmp_eff_deg_withlegend.png'), plot = eff_deg_comp,width=6,height=4)

eff_eff_comp <- ggplot(effector_pcc %>% mutate(go_group = fct_reorder(go_group, eff)) %>%
  mutate(eff_norm = eff / 5183), aes(x = go_group, y = eff_norm, fill = go_group)) +
  geom_boxplot() +
  # scale_x_discrete(labels=c('Chromosome\nSegregetion','Golgi vesicle\nTransport','Respiratory complex\nassembly'))+
  xlab("") +
  ggpubr::stat_compare_means(method = "kruskal.test") +
  scale_fill_manual(values = effector_colors) +
  ylab("Effectiveness") +
  theme_bw() +
  theme(
        axis.text.y=element_text(size=10),
        axis.title.y=element_text(size=14),axis.text.x = element_blank(), legend.title = element_blank())
ggsave(glue('{figure_path}//tmp_eff_eff.png'), plot = eff_eff_comp+theme(legend.position='none'),width=4,height=4)

sens_deg_comp <- ggplot(sensors_pcc %>% drop_na() %>% right_join(py$e_pcc$df)%>% mutate(go_group = factor(replace_na(as.character(go_group),"Non-sensor genes")),go_group=fct_reorder(go_group, sens),go_group=fct_relevel(go_group,'Non-sensor genes',after=Inf)), aes(x = go_group, y = deg, fill = go_group)) +
  geom_boxplot(
    outlier.shape=NA) +
  #    scale_x_discrete(labels=c('Chromosome\nSegregetion','Golgi vesicle\nTransport','Respiratory complex\nassembly'))+
  xlab("") +
  ylab("Degree") +
  scale_fill_manual(values = c(sensor_colors[1:6],'black')) +
  ggpubr::stat_compare_means(method = "kruskal.test", label.x.npc = .3) +
  theme_bw() +
  theme(
        axis.title.y=element_text(size=14),
        axis.text.y=element_text(size=10),
    axis.text.x = element_blank(), legend.title = element_blank(),
    legend.text = element_text(margin = margin(t = 10, b = 10))
  )+scale_y_log10()
  ggsave(glue('{figure_path}//sens_deg_withlegend.png'), plot=sens_deg_comp+theme(legend.position='bottom'),width=6,height=4)
ggsave(glue('{figure_path}//sens_deg.png'), plot=sens_deg_comp+theme(legend.position='none'),width=4,height=4)

sensors_pcc$go_group <- as.factor(sensors_pcc$go_group)
sensors_sens_pval <- FSA::dunnTest(sens ~ go_group, sensors_pcc)
sensors_sens_pval$res[sensors_sens_pval$res$P.adj < 0.1, ]

sens_sens_comp <- ggplot(sensors_pcc %>% drop_na() %>% mutate(go_group = fct_reorder(go_group, sens)) %>%
  mutate(sens_norm = sens / 5183), aes(x = go_group, y = sens_norm, fill = go_group)) +
  geom_boxplot() +
  geom_signif(
    comparisons = list(
      c("hexose metabolic\nprocess", "phenylalanine\ntransport"), c("tricarboxylic\nacid cycle", "phenylalanine\ntransport"),
      c("phenylalanine\ntransport", "protein folding")
    ),
    annotation = c("**", "**", "*"), y_position = c(0.007, 0.0075, 0.008)
  ) +
  #    scale_x_discrete(labels=c('Chromosome\nSegregetion','Golgi vesicle\nTransport','Respiratory complex\nassembly'))+
  xlab("") +
  ylab("Sensitivity") +
  scale_fill_manual(values = sensor_colors) +
  theme_bw() +
  theme(
        axis.title.y=element_text(size=14),
        axis.text.y=element_text(size=10),
        axis.text.x = element_blank(), legend.title = element_blank())
ggsave(glue('{figure_path}//sens_sens.png'), plot=sens_sens_comp+theme(legend.position='none'),width=4,height=4)

img <- "tmp_nw.png" %>%
  magick::image_read() # %>%
# magick::image_resize("570x380")

nw_img <- cowplot::ggdraw() +
  cowplot::draw_image(img, width = 1)
go_plots <- cowplot::plot_grid(
  sens_deg_comp + theme(legend.position = "none"),
  sens_sens_comp + theme(legend.position = "none"),
  cowplot::get_legend(sens_deg_comp), eff_deg_comp + theme(legend.position = "none"),
  eff_eff_comp + theme(legend.position = "none"),
  cowplot::get_legend(eff_deg_comp),
  ncol = 3, labels = c("B", "C", "", "D", "E", "")
)
combined <- cowplot::plot_grid(nw_img, go_plots, nrow = 2, rel_heights = c(0.4, 0.6), labels = c("A", ""))
ggsave(glue('{figure_path}//sens_deg.png'), plot=sens_deg_comp+theme(legend.position='none'),width=4,height=4)
ggsave(glue('{figure_path}//sens_sens.png'), plot=sens_sens_comp+theme(legend.position='none'),width=4,height=4)
#ggsave("reports/figures/paper_figures_0202_pcc_only/fig6.png", plot = combined, width = 9, height = 12)
```