### 2021-05-11 notebooks/CSHL_plots_022621.Rmd
more paper figures

---
title: "Paper figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
# use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv("enm")
setwd("../")
source("src/R/plot_density_prs.R")
source("src/R/get.subclusters.dendrogram.R")
```

```{python}
import matplotlib.pyplot as plt
import matplotlib as mpl
import networkx as nx
import numpy as np
import pandas as pd
import plotnine as p9
import pickle 
import os
import re
import itertools as itr
from src.enm import Enm

with open(f'data/interim/pcc_0909/pcc.pickle','rb') as f:
    e_pcc_old = pickle.load(f)
with open(f'data/interim/pcc_021521_withclustering/pcc_10_r.pickle','rb') as f:
    e_pcc = pickle.load(f)
pcc_df = e_pcc.df
pcc_gc = e_pcc.graph_gc
pcc_gc_pos = (dict(pcc_gc.nodes(data='pos')))
```

```{r}
library(tidyverse)
library(cowplot)
library(ggpubr)
library(ggsignif)
library(glue)
figure_folder = 'reports/figures/CSHL_plots_022621/'
pcc_df_random <- data.table::rbindlist(lapply(py$e_pcc$e_list, function(x) x$df), fill = T, idcol = "random_id") %>% mutate(random_id = as.factor(random_id))
pcc_df <- py$pcc_df %>%
  mutate(data = "pcc") %>%
  mutate(sens_norm = sens / nrow(.), eff_norm = eff / nrow(.), group = "real") %>%
  bind_rows(
    pcc_df_random %>% mutate(data = "pcc") %>% mutate(sens_norm = sens / nrow(py$pcc_df), eff_norm = eff / nrow(py$pcc_df), group = "random")
  ) %>%
  mutate(group = fct_relevel(group, c("real", "random")))
combined_df <- pcc_df
```

Degree - Eff - Sens distributions
```{r}

pcc_eff_dens <-
  plot_density_prs(pcc_df, eff_norm, xlab = "Effectiveness")
dens_lgn <- cowplot::get_legend(pcc_eff_dens + theme(legend.position = "right"))
pcc_degree_dens <- cowplot::ggdraw(plot_density_prs(pcc_df, deg, xlab = "Degree") + theme(legend.position = "none")) #+
#  cowplot::draw_plot(dens_lgn, .55,.55,.5,.5)
pcc_eff_dens <-
  cowplot::ggdraw(plot_density_prs(pcc_df, eff_norm, xlab = "Effectiveness") + theme(legend.position = "none") +
    coord_trans(x = "log10", xlim = c(NA, 10e-2)) +
    scale_x_continuous(
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    )) #+
#   cowplot::draw_plot(dens_lgn, .55,.55,.5,.5)
# annotate("text", x = 0.000001, y = 0.2, label=sprintf("p = %#.2f", eff_p), fontface='italic', size = 6)
pcc_sensitivity_dens <-
  cowplot::ggdraw(plot_density_prs(pcc_df, sens_norm, xlab = "Sensitivity") + theme(legend.position = "none")) #+
#        cowplot::draw_plot(dens_lgn, .55,.55,.5,.5)
# title <- cowplot::ggdraw() + cowplot::draw_label("Yeast GI Profile Similarity", fontface='bold')
pcc_dens_comb <- cowplot::plot_grid(pcc_eff_dens + theme(legend.position = "none"),
  pcc_sensitivity_dens + theme(legend.position = "none"),
  #+theme(legend.position='none'),
  ncol = 2
)
# pcc_dens_comb_title = cowplot::plot_grid(title, pcc_dens_comb, nrow=2 ,rel_heights=c(0.1,0.9))
# dens_lgn = cowplot::get_legend(pcc_sensitivity_dens)
dens_plots <- pcc_dens_comb # cowplot::plot_grid(pcc_dens_comb_title,nrow=1)
# cowplot::plot_grid(dens_plots, dens_lgn,nrow=2,rel_heights=c(.9,.1))
ggsave(glue('{figure_folder}/dens_comb.png', plot=pcc_dens_comb,width=8,height=8)
```

```{r}
real_random_labels <- c("PSN", "Rewired\nnetwork")
names(real_random_labels) <- c("real", "random")
pcc_deg_eff <- pcc_df %>% filter(group == "real") %>% # pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
  ggplot(aes(x = deg, y = eff_norm)) +
  geom_point() +
  #    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
  theme_bw() +
  ylab("Effectiveness") +
  xlab("Degree") +
  stat_cor(method = "pearson", cor.coef.name = "R", label.x.npc = 0.1) +
  theme(strip.text.y = element_text(angle = 0)) +
  scale_y_continuous(labels = scales::label_scientific(digits = 1))
pcc_deg_sens <- pcc_df %>% filter(group == "real") %>% # pivot_longer(cols=sens_norm:eff_norm, names_to="prs_data_type", values_to="prs_data_value")%>%
  ggplot(aes(x = deg, y = sens_norm)) +
  geom_point() +
  #    facet_grid(rows = vars(group), scales='free', labeller =labeller(group= real_random_labels ) ) +
  theme_bw() +
  ylab("Sensitivity") +
  xlab("Degree") +
  stat_cor(method = "spearman", cor.coef.name = "rho", label.x.npc = 0.1) +
  theme(strip.text.y = element_text(angle = 0)) +
  scale_y_continuous(labels = scales::label_scientific(digits = 1))
deg_scatter_plots <- cowplot::plot_grid(pcc_deg_eff, pcc_deg_sens,
#  labels = c("A", "B"),
  ncol = 2
)
ggsave(glue("{figure_folder}/fig2.png"), plot = deg_scatter_plots, width = 10, height = 5)
```

```{r}
pcc_rewire_df <- read_csv("data/interim/pcc_021521_withclustering//rewire_data.csv")
rewire_df <- pcc_rewire_df %>% mutate(data = "pcc")
network_labels <- c("Yeast GI Profile Similarity")
facet_labels <- network_labels
names(facet_labels) <- c("pcc")
combined_rewire_df <- rewire_df %>%
  select(data, eff_deg_pearson, sens_deg_spearman) %>%
  mutate(type = "random") %>%
  bind_rows(
    data.frame(
      sens_deg_spearman = c(cor(pcc_df[pcc_df$group == "real", ]$deg, pcc_df[pcc_df$group == "real", ]$sens, method = "spearman")),
      eff_deg_pearson = c(cor(pcc_df[pcc_df$group == "real", ]$deg, pcc_df[pcc_df$group == "real", ]$eff, method = "pearson")), type = "real", data = c("pcc")
    )
  ) %>%
  mutate(data = fct_relevel(data, c("pcc")))

p3 <-
  ggplot() +
  #    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
  #    geom_boxplot()+
  #    geom_jitter()+
  geom_density(data = combined_rewire_df %>% filter(type != "real"), aes(x = eff_deg_pearson, color = type, y = ..scaled..), linetype = "dashed") +
  geom_vline(data = combined_rewire_df %>% filter(type == "real"), aes(xintercept = eff_deg_pearson, color = type)) +
  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
  ylab("Density") +
  xlab("Effectiveness degree correlation") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines")) +
  scale_y_continuous(expand = c(0, 0))
p4 <-
  ggplot() +
      facet_grid(cols = vars(name), labeller = labeller(data=facet_labels))+
  #    geom_boxplot()+
  #    geom_jitter()+
  geom_boxplot(data = combined_rewire_df %>% filter(type != "real")%>%select(data, sens_deg_spearman,eff_deg_pearson)%>%pivot_longer(-data), 
               aes(y = value, x = name, color = name)) +
#  geom_density(data = combined_rewire_df %>% filter(type != "real"), aes(x = sens_deg_spearman, y = ..scaled.., color = type), linetype = "dashed") +
  geom_point(data = combined_rewire_df %>% filter(type == "real")%>%select(data, sens_deg_spearman,eff_deg_pearson)%>%pivot_longer(-data), 
             aes(x=name,y=value,color=name)) +
#  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
#  ylab("Density") +
#  xlab("Sensitivity degree correlation") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines"), legend.position = "bottom") 
#  scale_color_discrete(labels = c("Random networks", "Genetic network")) +
#  scale_y_continuous(expand = c(0, 0))
lgn34 <- cowplot::get_legend(p4)
distr_plots <- plot_grid(pcc_degree_dens + theme(legend.position = ""),
  ggdraw(p3 + theme(legend.position = "")),
  ggdraw(p4 + theme(legend.position = "")),
#  labels = c("C", "D", "E"), 
  nrow = 3
)
# distr_plots_withlgn <- plot_grid(distr_plots, lgn34,
# nrow=2,rel_heights = c(0.9,0.1))
# ggsave('reports/figures/paper_figures_0916/fig2.png',plot=distr_plots , width=7)
ggsave(glue("{figure_folder}/distr_plots.pdf"), plot= p4, width = 4, height = 10)

#plot_grid(deg_scatter_plots, distr_plots, dens_plots, nrow = 3, rel_heights = c(0.3, 0.4, .3))
#ggsave(glue("{figure_folder}fig3.pdf"), width = 10, height = 12)
```

```{r}
library(RColorBrewer)
col <- brewer.pal(n = 8, name = "Set1")
sensor_colors <- c(
'#dadaeb',
'#bcbddc',
'#9e9ac8',
'#807dba',
'#6a51a3',
'#4a1486'
)#col[1:4] # brewer.pal(n = 6, name = "Set1")
effector_colors <- c('#bae4b3',
'#74c476',
'#238b45')#col[5:8] # brewer.pal(n=3, name= "Set1_r")
```

```{python}
sensor_colors = r.sensor_colors
effector_colors = r.effector_colors
sensors_pcc = e_pcc_old.sensors_df
sensors_pcc['go_group'] = sensors_pcc['cluster_go_term']
sensors_pcc['go_group'] = ['None'  if pd.isna(i) else i for i in sensors_pcc['go_group'].tolist()]
effector_pcc = e_pcc.effectors_df
effector_pcc['go_group'] = effector_pcc['cluster_go_term']
effector_order = effector_pcc.groupby('go_group').clusters.median().sort_values().index.tolist()
sensor_order = sensors_pcc.groupby('go_group').clusters.median().sort_values().index.tolist()
#sensor_order.append(None)
```
# Dendrogram cluster network plots
```{python}
import copy
import scipy.cluster.hierarchy as sch
prs_mat = e_pcc.prs_mat
quantile_threshold =  0.95
method = 'ward'
metric = 'seuclidean'
q99 = np.quantile(prs_mat, quantile_threshold)
prs_mat_cl = copy.deepcopy(prs_mat)
prs_mat_cl[prs_mat_cl > q99] = q99
prs_mat_cl_orig = prs_mat_cl
# , optimal_ordering=True)
#row_linkage_cl = sch.linkage(sch.distance.pdist(prs_mat_cl), method=method)
# ,optimal_ordering=True)
#col_linkage_cl = sch.linkage(sch.distance.pdist(prs_mat_cl.T), method=method)

row_dist_orig = sch.distance.pdist((prs_mat),metric = metric)
col_dist_orig = sch.distance.pdist((prs_mat.T),metric = metric)
row_dist_orig_sqr = sch.distance.squareform(row_dist_orig)
col_dist_orig_sqr = sch.distance.squareform(col_dist_orig)

row_dist = sch.distance.pdist(prs_mat_cl,metric = metric)
col_dist = sch.distance.pdist(prs_mat_cl.T,metric = metric)

row_dist_sqr = sch.distance.squareform(row_dist)
col_dist_sqr = sch.distance.squareform(col_dist)

```

```{r}
cluster_method = 'ward.D2'
dist_row = as.dist(py$row_dist_sqr)
hclust_row = hclust(dist_row, cluster_method)
dist_col = as.dist(py$col_dist_sqr)
hclust_col = hclust(dist_col, cluster_method)
dend_row  <- as.dendrogram(hclust_row)
dend_col  <- as.dendrogram(hclust_col)
#effector_pcc <- get.subclusters.dendrogram(dend_row, effector_pcc)
#sensors_pcc <- get.subclusters.dendrogram(dend_col, sensors_pcc%>%select(-rowname))
library(dendextend)
png(glue('{figure_folder}/row_dendrogram_colored.png'),width=400,height=400)
dend_row_plot <- dend_row[[2]] %>% 
    set("branches_k_color", k = 4,value = effector_colors[c(4,2,3,1)]) %>% 
    set("labels", NA) %>%    # change labelsk
    as.ggdend()%>%
    ggplot()+
    coord_flip()#    plot(xlab="", sub="")}
dend_row_plot
dev.off()
png(glue('{figure_folder}/col_dendrogram_colored.png'),width=400,height=400)
dend_col_plot <-     dend_col[[1]] %>% 
    set("branches_k_color", k = 4,value = sensor_colors[c(3,2,4,1)]) %>% 
    set("labels", NA) %>%    # change labelsk
    as.ggdend()%>%
    ggplot()
dend_col_plot
dev.off()
```

```{python}
from matplotlib.lines import Line2D
pos = e_pcc_old.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#                         node_shape=['^' if i in sensors_pcc.orf_name.values else 'v' if i in effector_pcc.orf_name.values else 'o' for i in e_pcc.nodes])
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=sensors_pcc.orf_name.values, node_size=200, pos=pos,
                          node_color='black',
                          node_shape='^',edgecolors='black',
                          linewidths=1)
for itr, i in enumerate(sensor_order):
    orf_names_to_plot = sensors_pcc.loc[sensors_pcc.go_group==i, 'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=sensor_colors[itr],
                          node_shape='^',edgecolors='black',
                          linewidths=1)
    legend_elements.append(
        Line2D([0], [0], marker='^', color='black', label=f'Column cluster ({itr+1})',
                              markerfacecolor=sensor_colors[itr], markersize=12, linestyle="None")
    )
#nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, sensors_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='red', alpha=0.5)
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-")]
#                   Line2D([0], [0], marker='o', color='red', label= 'Sensor-Sensor edges',
#                              markerfacecolor='#018571', markersize=0, linestyle="-")]
)
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))
plt.savefig(f'{r.figure_folder}tmp_nw_sens.png',bbox_inches='tight')

```

```{python}
from matplotlib.lines import Line2D
pos = e_pcc_old.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#                         node_shape=['^' if i in sensors_pcc.orf_name.values else 'v' if i in effector_pcc.orf_name.values else 'o' for i in e_pcc.nodes])
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=effector_pcc.orf_name.values, node_size=100, pos=pos,
                      node_color='black',
                      node_shape='o')
for itr, i in enumerate(effector_order):
    orf_names_to_plot = effector_pcc.loc[effector_pcc.go_group==i,'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=effector_colors[itr],
                          node_shape='o',edgecolors='black',
                          linewidths=1)
    legend_elements.append(
            Line2D([0], [0], marker='o', color='black', label=f'Row cluster ({itr+1})',
                                  markerfacecolor=effector_colors[itr], markersize=12, linestyle="None")
        )
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-")]
#                   Line2D([0], [0], marker='o', color='blue', label= 'Effector-Effector edges',
 #                             markerfacecolor='#a6611a', markersize=0, linestyle="-")]
)
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))

#nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, effector_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='blue',alpha=0.5)
plt.savefig(f'{r.figure_folder}tmp_nw_eff.png',bbox_inches='tight')

from src.visualize.visualize import heatmap_annotated
#import src
#import importlib
#importlib.reload(src.visualize.visualize)
heatmap_annotated(prs_mat, prs_mat_cl_orig, figure_path=r.figure_folder, row_linkage=e_pcc.row_linkage, col_linkage=e_pcc.col_linkage, save_figure=True)#,
#                 row_colors=colors, col_colors=colors)

```

```{r}
library(coin)
df <- py$e_pcc$df%>%mutate(eff_group = as.factor(ifelse(orf_name%in%effector_pcc$orf_name,'eff','other')),
                    sens_group = factor(ifelse(orf_name%in%sensors_pcc$orf_name,'sens','other'),levels=c('sens','other')),
                    eff_norm = eff / nrow(.),
                    sens_norm = sens/ nrow(.))
pval_eff = pvalue(independence_test(eff~eff_group,df))
pval_sens = pvalue(independence_test(sens~sens_group,df))

eff_group_box <- df%>%ggplot(aes(x=eff_group,y=eff_norm))+
    geom_boxplot()+
  geom_signif(comparisons = list(c("eff", "other")), 
              map_signif_level=TRUE, y_position = 0.006)+
    xlab('')+
    ylab('Effectiveness')+
    scale_x_discrete(labels=c('Effectors','Others'))+
    theme_bw()
sens_group_box <- df%>%ggplot(aes(x=sens_group,y=sens_norm))+
    geom_boxplot()+
  geom_signif(comparisons = list(c("sens", "other")), 
              map_signif_level=TRUE, y_position = 0.0075)+
    xlab('')+
    ylab('Sensitivity')+
    scale_x_discrete(labels=c('Sensors','Others'))+
    theme_bw()+ylim(0,0.008)
#ggsave('tmp.png',sens_group_box)
heatmap_img <- glue('{figure_folder}prs_heatmap.png')%>%
    magick::image_read()
heatmap_img <- cowplot::ggdraw()+
    cowplot::draw_image(heatmap_img, width = 1)
eff_nw_img <- glue('{figure_folder}tmp_nw_eff.png') %>%
  magick::image_read() # %>%
eff_nw_img <- cowplot::ggdraw() +
  cowplot::draw_image(eff_nw_img, width = 1)
sens_nw_img <- glue('{figure_folder}tmp_nw_sens.png') %>%
  magick::image_read() # %>%
sens_nw_img <- cowplot::ggdraw() +
  cowplot::draw_image(sens_nw_img, width = 1)
cowplot::plot_grid(plot_grid(heatmap_img, plot_grid(sens_group_box, eff_group_box+ylim(0,0.007), nrow=2,labels=c('B','C')), nrow=1, rel_widths = c(0.7,0.3),labels=c('A','')),
                   plot_grid(dend_col_plot, sens_nw_img, nrow=1, rel_widths=c(0.2,.8), labels=c('D','E')),
                   plot_grid(dend_row_plot, eff_nw_img, nrow=1, rel_widths=c(0.2,.8), labels=c('F','G')), nrow=3, rel_heights=c(.5,.25,.25))
ggsave(glue('{figure_folder}fig1.pdf'),height=10,width=8)

```

# Go term network plots
```{python}
from matplotlib.lines import Line2D
pos = e_pcc_old.graph_gc.nodes('pos')
fig, ax = plt.subplots(figsize=(12,12))
legend_elements = [    ]
e_pcc.plot_network_spring(ax=ax,
                          node_size=1,
                          node_color='black',
 #                        node_size = [100 if i in sensors_pcc.orf_name.values or i in effector_pcc.orf_name.values else 1 for i in e_pcc.nodes],
                         #node_color = ['red' if i in sensors_pcc.orf_name.values else 'blue' if i in effector_pcc.orf_name.values else 'black' for i in e_pcc.nodes],
                         edge_color='black')
#                         node_shape=['^' if i in sensors_pcc.orf_name.values else 'v' if i in effector_pcc.orf_name.values else 'o' for i in e_pcc.nodes])
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=sensors_pcc.orf_name.values, node_size=200, pos=pos,
                          node_color='black',
                          node_shape='^',edgecolors='black',
                          linewidths=1)
for itr, i in enumerate(sensor_order):
    #print(i, effector_colors[itr])

    orf_names_to_plot = sensors_pcc.loc[sensors_pcc.go_group==i, 'orf_name'].tolist()
#    print(orf_names_to_plot)
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=sensor_colors[itr],
                          node_shape='^',edgecolors='black',
                          linewidths=1)
    legend_elements.append(
        Line2D([0], [0], marker='^', color='black', label=f'Column Cluster ({i})',
                              markerfacecolor=sensor_colors[itr], markersize=12, linestyle="None")
    )
nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=effector_pcc.orf_name.values, node_size=100, pos=pos,
                      node_color='black',
                      node_shape='o')
for itr, i in enumerate(effector_order):
    orf_names_to_plot = effector_pcc.loc[effector_pcc.go_group==i,'orf_name'].tolist()
    nx.draw_networkx_nodes(e_pcc.graph_gc, nodelist=orf_names_to_plot, node_size=200, pos=pos,
                          node_color=effector_colors[itr],
                          node_shape='o',edgecolors='black',
                          linewidths=1)
    if i == 'mito':
        lbl = 'Respiratory complex assembly'
    elif i == 'golgi':
        lbl = "Golgi vesicle transport"
    else:
        lbl = 'Chromosome segregation'
    legend_elements.append(
            Line2D([0], [0], marker='o', color='black', label=f'Effectors ({i})',
                                  markerfacecolor=effector_colors[itr], markersize=12, linestyle="None")
        )
nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, sensors_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='red', alpha=0.5)
ax.set_facecolor('white')
legend_elements.extend(
    [Line2D([0], [0], marker='o', color='black', label='Genes',
                              markerfacecolor='black', markersize=4, linestyle="None"),
#                    Line2D([0], [0], marker='o', color='black', label='Effectors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
#                    Line2D([0], [0], marker='^', color='black', label='Sensors',
#                               markerfacecolor='black', markersize=10, linestyle="None"),
                       Line2D([0], [0], marker='o', color='black', label= 'High functional similarity',
                              markerfacecolor='black', markersize=0, linestyle="-"),
                   Line2D([0], [0], marker='o', color='red', label= 'Sensor-Sensor edges',
                              markerfacecolor='#018571', markersize=0, linestyle="-"),
                   Line2D([0], [0], marker='o', color='blue', label= 'Effector-Effector edges',
                              markerfacecolor='#a6611a', markersize=0, linestyle="-")]
)
lgd = ax.legend(handles=legend_elements, fontsize=22,loc='center left', bbox_to_anchor=(1.1, 0.5))

nx.draw_networkx_edges(nx.induced_subgraph(e_pcc.graph_gc, effector_pcc.orf_name.tolist()), ax=ax , pos=pos, edge_color='blue',alpha=0.5)
plt.savefig('tmp_nw.png',bbox_inches='tight')

```

```{r}
effector_pcc <- py$effector_pcc
sensors_pcc <- py$sensors_pcc %>% mutate(go_group = (as.character(go_group)), go_group = ifelse(go_group == "NaN", NA, go_group))%>%drop_na()
# library(tidyverse)
effector_pcc$go_group <- as.factor(effector_pcc$go_group)

# res.aov <- kruskal.test(eff~go_group, data= effector_pcc)
# res.aov
effector_eff_pval <- FSA::dunnTest(eff ~ go_group, effector_pcc)
effector_eff_pval$res[effector_eff_pval$res$P.adj < 0.1, ]
effector_deg_pval <- FSA::dunnTest(deg ~ go_group, effector_pcc)
effector_deg_pval$res[effector_deg_pval$res$P.adj < 0.1, ]

eff_deg_comp <- ggplot(effector_pcc %>% mutate(go_group = fct_reorder(go_group, eff)), aes(x = go_group, y = deg, fill = go_group)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(
    values = effector_colors,
#    labels = c("Chromosome\nSegragetion", "Golgi vesicle\nTransport", "Respiratory complex\nassembly")
  ) +
  xlab("") +
  ylab("Degree") +
  theme_bw() +
  geom_signif(
    comparisons = list(c("golgi", "chromatin"), c("golgi", "mito"), c("mito", "chromatin")),
    annotation = c("**", "***", "ns"),
    y_position = c(110, 115, 120)
  ) +
  theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.text = element_text(margin = margin(t = 10, b = 10))) # , legend.spacing.y = unit(2,'in'))+
eff_eff_comp <- ggplot(effector_pcc %>% mutate(go_group = fct_reorder(go_group, eff)) %>%
  mutate(eff_norm = eff / 5183), aes(x = go_group, y = eff_norm, fill = go_group)) +
  geom_boxplot() +
  # scale_x_discrete(labels=c('Chromosome\nSegragetion','Golgi vesicle\nTransport','Respiratory complex\nassembly'))+
  xlab("") +
  ggpubr::stat_compare_means(method = "kruskal.test") +
  scale_fill_manual(values = effector_colors) +
  ylab("Effectiveness") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.title = element_blank())
ggsave("tmp.png", plot = eff_eff_comp)

sens_deg_comp <- ggplot(sensors_pcc %>% drop_na() %>% mutate(go_group = fct_reorder(go_group, sens)), aes(x = go_group, y = deg, fill = go_group)) +
  geom_boxplot() +
  #    scale_x_discrete(labels=c('Chromosome\nSegragetion','Golgi vesicle\nTransport','Respiratory complex\nassembly'))+
  xlab("") +
  ylab("Degree") +
  scale_fill_manual(values = sensor_colors) +
  ggpubr::stat_compare_means(method = "kruskal.test", label.x.npc = .3) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(), legend.title = element_blank(),
    legend.text = element_text(margin = margin(t = 10, b = 10))
  ) # , legend.spacing.y = unit(2,'cm'))
sensors_pcc$go_group <- as.factor(sensors_pcc$go_group)
sensors_sens_pval <- FSA::dunnTest(sens ~ go_group, sensors_pcc)
sensors_sens_pval$res[sensors_sens_pval$res$P.adj < 0.1, ]

sens_sens_comp <- ggplot(sensors_pcc %>% drop_na() %>% mutate(go_group = fct_reorder(go_group, sens)) %>%
  mutate(sens_norm = sens / 5183), aes(x = go_group, y = sens_norm, fill = go_group)) +
  geom_boxplot() +
  ggpubr::stat_compare_means(method = "kruskal.test", label.x.npc = .3) +
#  geom_signif(
#    comparisons = list(
#      c("hexose metabolic\nprocess", "phenylalanine\ntransport"), c("tricarboxylic\nacid cycle", "phenylalanine\ntransport"),
#      c("phenylalanine\ntransport", "protein folding")
#    ),
#    annotation = c("**", "**", "*"), y_position = c(0.007, 0.0075, 0.008)
#  ) +
  #    scale_x_discrete(labels=c('Chromosome\nSegragetion','Golgi vesicle\nTransport','Respiratory complex\nassembly'))+
  xlab("") +
  ylab("Sensitivity") +
  scale_fill_manual(values = sensor_colors) +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.title = element_blank())

img <- "tmp_nw.png" %>%
  magick::image_read() # %>%
# magick::image_resize("570x380")

nw_img <- cowplot::ggdraw() +
  cowplot::draw_image(img, width = 1)
go_plots <- cowplot::plot_grid(
  sens_deg_comp + theme(legend.position = "none"),
  sens_sens_comp + theme(legend.position = "none"),
  cowplot::get_legend(sens_deg_comp), eff_deg_comp + theme(legend.position = "none"),
  eff_eff_comp + theme(legend.position = "none"),
  cowplot::get_legend(eff_deg_comp),
  ncol = 3, labels = c("B", "C", "", "D", "E", "")
)
combined <- cowplot::plot_grid(nw_img, go_plots, nrow = 2, rel_heights = c(0.4, 0.6), labels = c("A", ""))
ggsave(glue("{figure_folder}/fig6.png"), plot = combined, width = 9, height = 12)

```