### 2021-05-19 notebooks/25-Figure3-051821.Rmd

---
title: "25-Figure3-051821"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(glue)
library(cowplot)
library(ggpubr)
library(ggsignif)
source("../enm/R/plot_density_prs.R")#paste function in this file
```

# Sensor connectivity and neighbor degree

```{r}
sensor_connectivity_df <- read_csv("../data/interim/sensor_connectivity_df.csv") %>%
  #mutate(data = fct_relevel(data, c("pcc", "gi", "yuri", "huri"))) %>%
  filter(data == "pcc")
pcc_df <- read_csv('../data/interim/pcc_df.csv')
# effector_connectivity_df <- read_csv('data/interim/sensor_connectivity_1205/effector_connectivity_df.csv')%>%
#    mutate(data= fct_relevel(data, c('pcc','gi','yuri','huri')))%>%filter(data=='pcc')
# effector_connectivity_plot <-
#    ggplot()+
##    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
##    geom_boxplot()+
##    geom_jitter()+
#    geom_density(data =effector_connectivity_df%>%filter(type!='real'),aes(x=ratio,color=type, y=..scaled..),linetype = "dashed")+
#    geom_vline(data =effector_connectivity_df%>%filter(type=='real'),aes(xintercept=ratio,color=type))+
#    guides(color=guide_legend(title="Data type"),shape=guide_legend(title="Network type"))+
#    ylab('Density')+xlab('Ratio of between group edges to outgroup edges')+
#    theme_bw()+
#    scale_color_discrete(labels=c('Random sampled set', 'Effectors'))+
#    theme(legend.position='bottom')+
#    scale_y_continuous(expand = c(0, 0))
eff_neighbor_degree_plot <- pcc_df %>%
  #filter(group == "real") %>%
  #group_by(data) %>%
  mutate(sens_norm = sens / nrow(.), eff_norm = eff / nrow(.), group = "real")%>%
  mutate(is_effector = factor(ifelse(eff_norm > quantile(eff_norm, 0.99), TRUE, FALSE), levels = c(TRUE, FALSE))) %>%
  #    mutate(is_effector = fct_relevel(is_effector,c(TRUE,FALSE)))%>%
  ggplot(aes(x = is_effector, y = log(neighbor_degree))) +
  #    facet_grid(cols=vars(data),labeller = labeller(data=facet_labels))+
  #    geom_point()+
  geom_boxplot() +
  theme_bw() +
  ylab("Log(Average neighbor degree)") +
  xlab("") +
  stat_compare_means() +
  scale_x_discrete(labels = c("Effector", "Not Effector"))
sensor_connectivity_plot <-
  ggplot() +
  #    facet_grid(cols = vars(data), labeller = labeller(data=facet_labels))+
  #    geom_boxplot()+
  #    geom_jitter()+
  geom_density(data = sensor_connectivity_df %>% filter(type != "real"), aes(x = ratio, color = type)) +
  geom_vline(data = sensor_connectivity_df %>% filter(type == "real"), aes(xintercept = ratio, color = type)) +
  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
  ylab("Density") +
  xlab("Ratio of between group edges\nto outgroup edges") +
  theme_bw() +
  scale_color_discrete(labels = c("Random sampled set", "Sensors")) +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0, 0))
sensor_connectivity_plot
neighbor_degree_plot <- combined_df %>%
  filter(group == "real") %>%
  group_by(data) %>%
  mutate(is_sensor = factor(ifelse(sens_norm > quantile(sens_norm, 0.99), TRUE, FALSE), levels = c(TRUE, FALSE))) %>%
  #    mutate(is_sensor = fct_relevel(is_sensor,c(TRUE,FALSE)))%>%
  ggplot(aes(x = is_sensor, y = log(neighbor_degree))) +
  #    facet_grid(cols=vars(data),labeller = labeller(data=facet_labels))+
  #    geom_point()+
  geom_boxplot() +
  theme_bw() +
  ylab("Log(Average neighbor degree)") +
  xlab("") +
  stat_compare_means() +
  scale_x_discrete(labels = c("Sensor", "Not Sensor"))
sensor_conn_lgn <- get_legend(sensor_connectivity_plot + theme(legend.position = "right"))
plot_grid(neighbor_degree_plot,
  sensor_connectivity_plot + theme(legend.position = "none"), # effector_connectivity_plot,
  eff_neighbor_degree_plot, sensor_conn_lgn,
  nrow = 2,
  labels = c("A", "B", "C", "")
) # %>%
#    plot_grid(, labels=c('','C'),nrow=2, rel_widths=c(2,1))
#ggsave("reports/figures/paper_figures_0202_pcc_only/fig5.pdf", width = 6, height = 6)
sensor_connectivity_df$ratio[sensor_connectivity_df$type=='random']/mean(sensor_connectivity_df$ratio[sensor_connectivity_df$type=='random'])
```