### 2021-05-24 notebooks/03-Fig3abde_4acd_5b-051821.Rmd
recover pcc.py and edit plot legend labels

---
title: "25-Figure3-051821"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

# Load packages
```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(glue)
library(cowplot)
library(ggpubr)
library(ggsignif)
library(RColorBrewer)
col <- brewer.pal(n = 9, name = "Set1")
sensor_colors <- col[2:6] # brewer.pal(n = 6, name = "Set1")
effector_colors <- col[7:9] # brewer.pal(n=3, name= "Set1_r")
if(!require("FSA")){
  remotes::install_github('droglenc/FSA')
}
```

```{r}
sessionInfo()
```
# Load necessary data
```{r}

sensor_connectivity_df <- read_csv(snakemake@input[['sensor_connectivity_df']])#"../data/interim/sensor_connectivity_df.csv") 
pcc_df <- read_csv(snakemake@input[['pcc_df']])#'../data/interim/pcc_df.csv')
sensors_pcc <- read_csv(snakemake@input[['sensors_pcc']])%>%#'../data/interim/sensors_df.csv')%>%
 mutate(go_group = (as.character(go_group)), go_group = ifelse(go_group == "NULL", NA, go_group))
effector_pcc <- read_csv(snakemake@input[['effector_pcc']])#'../data/interim/effectors_df.csv')

# Ranked groups GO enrichment results
rwr_row <- read_csv(snakemake@input[['rwr_row']])%>%mutate(data = 'rwr_row')
prs_row <- read_csv(snakemake@input[['prs_row']])%>%mutate(data = 'prs_row')
rwr_column <- read_csv(snakemake@input[['rwr_column']])%>%mutate(data = 'rwr_col')
prs_column <- read_csv(snakemake@input[['prs_column']])%>%mutate(data = 'prs_col')
go_combined <- data.table::rbindlist(list(rwr_row,prs_row,rwr_column,prs_column))
```
# Figure3
## A
```{r}
sensor_neighbor_degree_plot <- pcc_df %>%
  mutate(is_sensor = factor(ifelse(sens > quantile(sens, 0.99), TRUE, FALSE), levels = c(TRUE, FALSE))) %>%
  ggplot(aes(x = is_sensor, y = log(neighbor_degree))) +
  geom_boxplot() +
  theme_bw() +
  ylab("Log(Average neighbor degree)") +
  xlab("") +
  stat_compare_means() +
  scale_x_discrete(labels = c("Sensor", "Not Sensor"))
  sensor_neighbor_degree_plot

```

## B
```{r}
sensor_connectivity_plot <-
  ggplot() +
  geom_density(data = sensor_connectivity_df %>% filter(type != "real"), aes(x = ratio, color = type)) +
  geom_vline(data = sensor_connectivity_df %>% filter(type == "real"), aes(xintercept = ratio, color = type)) +
  guides(color = guide_legend(title = "Data type"), shape = guide_legend(title = "Network type")) +
  ylab("Density") +
  xlab("Ratio of between group edges\nto outgroup edges") +
  theme_bw() +
  scale_color_discrete(labels = c("Random sampled set", "Sensors")) +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0, 0))
  sensor_connectivity_plot
```

## D
```{r}

sens_deg_comp <- ggplot(sensors_pcc %>% drop_na() %>% right_join(pcc_df)%>% mutate(go_group = factor(replace_na(as.character(go_group),"Non-sensor genes")),go_group=fct_reorder(go_group, sens),go_group=fct_relevel(go_group,'Non-sensor genes',after=Inf)), aes(x = go_group, y = deg, fill = go_group)) +
  geom_boxplot(
    outlier.shape=NA) +
  xlab("") +
  ylab("Degree") +
  scale_fill_manual(values = c(sensor_colors[1:5],'black')) +
  ggpubr::stat_compare_means(method = "kruskal.test", label.x.npc = .3) +
  theme_bw() +
  theme(
        axis.title.y=element_text(size=14),
        axis.text.y=element_text(size=10),
    axis.text.x = element_blank(), legend.title = element_blank(),
    legend.text = element_text(margin = margin(t = 10, b = 10))
  )+scale_y_log10()
sens_deg_comp

```

## E
```{r}

sensors_pcc$go_group <- as.factor(sensors_pcc$go_group)%>%fct_recode(
  `Iron ion transport`  = "cellular response to iron ion starvation",
  `Mitochondria-nucleus\nsignaling pathway\nand\nTricarboxylic acid cycle`="mitochondria-nucleus signaling pathway",
  `Phenylalanine transport`="phenylalanine transport",
  `Hexose metabolic process`="hexose metabolic process",
  `Tricarboxylic acid cycle`="tricarboxylic acid cycle"
)
sensors_sens_pval <- FSA::dunnTest(sens ~ go_group, sensors_pcc)
sensors_sens_pval$res[sensors_sens_pval$res$P.adj < 0.1, ]

sens_sens_comp <- ggplot(sensors_pcc %>% drop_na() %>% mutate(go_group = fct_reorder(go_group, sens)) %>%
  mutate(sens_norm = sens / 5183), aes(x = go_group, y = sens_norm, fill = go_group)) +
  geom_boxplot() +
  geom_signif(
    comparisons = list(
      c("Hexose metabolic process", "Phenylalanine transport"), c("Tricarboxylic acid cycle", "Phenylalanine transport")#,
      #c("phenylalanine\ntransport", "protein folding")
    ),
    annotation = c("**", "**"), y_position = c(0.007, 0.0075)#, 0.008)
  ) +
  xlab("") +
  ylab("Sensitivity") +
  scale_fill_manual(values = sensor_colors) +
  theme_bw() +
  theme(
        axis.title.y=element_text(size=14),
        axis.text.y=element_text(size=10),
        axis.text.x = element_blank(),
        legend.title = element_blank())

sens_sens_comp
```

# Figure4
## A
```{r}
eff_neighbor_degree_plot <- pcc_df %>%
  mutate(sens_norm = sens / nrow(.), eff_norm = eff / nrow(.), group = "real")%>%
  mutate(is_effector = factor(ifelse(eff_norm > quantile(eff_norm, 0.99), TRUE, FALSE), levels = c(TRUE, FALSE))) %>%
  ggplot(aes(x = is_effector, y = log(neighbor_degree))) +
  geom_boxplot() +
  theme_bw() +
  ylab("Log(Average neighbor degree)") +
  xlab("") +
  stat_compare_means() +
  scale_x_discrete(labels = c("Effector", "Not Effector"))
eff_neighbor_degree_plot
```

## C
```{r}
effector_pcc$go_group <- as.factor(effector_pcc$go_group)%>%fct_recode(
  golgi = "phosphatidylinositol biosynthetic process",
  mito = "mitochondrial cytochrome c oxidase assembly",
  chromosome = "replication fork protection"
)

# res.aov <- kruskal.test(eff~go_group, data= effector_pcc)
# res.aov
effector_eff_pval <- FSA::dunnTest(eff ~ go_group, effector_pcc)
effector_eff_pval$res[effector_eff_pval$res$P.adj < 0.1, ]
effector_deg_pval <- FSA::dunnTest(deg ~ go_group, effector_pcc)
effector_deg_pval$res[effector_deg_pval$res$P.adj < 0.1, ]

eff_deg_comp <- ggplot(effector_pcc %>% right_join(pcc_df)%>% mutate(go_group = factor(replace_na(as.character(go_group),"Others")),go_group=fct_reorder(go_group, eff),go_group=fct_relevel(go_group,'Others',after=Inf)), 
                       aes(x = go_group, y = deg, fill = go_group)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(
    values = c(effector_colors,'black'),
    labels = c("Chromosome\nSegregetion", "Golgi vesicle\nTransport", "Respiratory complex\nassembly",'Non-effector genes')
  ) +
  xlab("") +
  ylab("Degree") +
  theme_bw() +
  geom_signif(
    comparisons = list(c("golgi", "chromosome"), c("golgi", "mito"), c("mito", "chromosome")),
    annotation = c("**", "***", "ns"),
    y_position = c(115, 120, 130)
  ) +
  theme(
       axis.text.y=element_text(size=10),
        axis.title.y=element_text(size=14),legend.title = element_blank(),  axis.text.x = element_blank(), legend.text = element_text(margin = margin(t = 10, b = 10))) # , legend.spacing.y = unit(2,'in'))+
eff_deg_comp
```

## D
```{r}

eff_eff_comp <- ggplot(effector_pcc %>% mutate(go_group = fct_reorder(go_group, eff)) %>%
  mutate(eff_norm = eff / 5183), aes(x = go_group, y = eff_norm, fill = go_group)) +
  geom_boxplot() +
  xlab("") +
  ggpubr::stat_compare_means(method = "kruskal.test") +
  scale_fill_manual(values = effector_colors,
  labels = c("Chromosome\nSegregetion", "Golgi vesicle\nTransport", "Respiratory complex\nassembly")) +
  ylab("Effectiveness") +
  theme_bw() +
  theme(
        axis.text.y=element_text(size=10),
        axis.title.y=element_text(size=14),axis.text.x = element_blank(), legend.title = element_blank())
eff_eff_comp
```

# Figure 5
## B
```{r}
ci <- function(p,n){
   sd = sqrt(p*(1-p)/n)
   ci = 1.96*sd
   ci
}
# New facet label names for dose variable
axis.labs <- c("Row-based", "Column-based")
names(axis.labs) <- c("row", "col")
go_summarised <- go_combined%>%
  group_by(data)%>%
  summarise(positives=n_distinct(orf_name_id))%>%
  mutate(n_tot=5183, prop = positives/n_tot, ci_min=prop-ci(prop,n_tot),ci_max=prop+ci(prop,n_tot))%>%
  separate(data,into=c('method','axis'))%>%
  mutate(axis=factor(axis,levels=c('row','col')))

p_val_row  <- go_summarised%>%
  mutate(negatives = n_tot-positives)%>%
  dplyr::filter(axis=="row")%>%
  select(positives,negatives)%>%
  as.matrix()%>%
  prop.test(correct=FALSE)
p_val_col <- go_summarised%>%
  mutate(negatives = n_tot-positives)%>%
  dplyr::filter(axis=="col")%>%
  select(positives,negatives)%>%
  as.matrix()%>%
  prop.test(correct=FALSE)

annotation_df <- data.frame(
  axis = factor(c("row", "col")),
  start = c("prs", "prs"),
  end = c("rwr", "rwr"),
  y = c(0.91, 0.91),
  label = c(glue::glue("p = {format.pval(p_val_row$p.value,digits=2)}"), glue::glue("p{format.pval(p_val_col$p.value)}"))
)

 ggplot()+
  facet_grid(vars(axis),
  labeller = labeller(axis=axis.labs))+
  geom_col(data=go_summarised,aes(x=method,y=prop,fill=method))+
  geom_errorbar(data=go_summarised, aes(x=method, y=prop, ymin=ci_min, ymax=ci_max), width=.2) +
  theme_bw()+
  scale_x_discrete(labels=c('PRS','RWR'))+
  xlab('')+
  ylab('%of tests with GO enrichments')+
  theme(
   strip.background = element_rect(
     color=,"black", fill="white", size=1.5, linetype="solid"
     ),
     legend.position='none',
     panel.spacing = unit(2, "lines")
   )+
   geom_signif(
    data = annotation_df,
    aes(xmin = start, xmax = end, annotations = label, y_position = y),
    textsize = 4,tip_length = .1,
    manual = TRUE
  )+
  scale_y_continuous(expand=c(0,0),limits=c(0,1))
```