### 2021-05-30 notebooks/04-Fig5b-052721.Rmd
install r dependencies using remotes, instead of conda

---
title: "04-Figure5b"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

# Load packages
```{r, message=FALSE,warning=FALSE}
if (!require('tidyverse')) {
   remotes::install_version("tidyverse", version = "1.3.0", repos = "http://cran.us.r-project.org")
   library(tidyverse)
}
if (!require('glue')) {
   remotes::install_version("glue", version = "1.4.1", repos = "http://cran.us.r-project.org")
   library(glue)
}
if (!require('cowplot')) {
   remotes::install_version("cowplot", version = "1.0.0", repos = "http://cran.us.r-project.org")
   library(cowplot)
}
if (!require('ggpubr')) {
   remotes::install_version('foreign',version="0.8-72", repos = "http://cran.us.r-project.org")
   remotes::install_version("ggpubr", version = "0.4.0", repos = "http://cran.us.r-project.org")
   library(ggpubr)
}
if (!require('ggsignif')) {
   remotes::install_version("ggsignif", version = "0.6.0", repos = "http://cran.us.r-project.org")
   library(ggsignif)
}
```

# Figure 5
## B
```{r}

# Ranked groups GO enrichment results
effector_pcc <- read_csv(snakemake@input[['effector_pcc']])#'../data/interim/effectors_df.csv')
rwr_row <- read_csv(snakemake@input[['rwr_row']])%>%mutate(data = 'rwr_row')
prs_row <- read_csv(snakemake@input[['prs_row']])%>%mutate(data = 'prs_row')
rwr_column <- read_csv(snakemake@input[['rwr_column']])%>%mutate(data = 'rwr_col')
prs_column <- read_csv(snakemake@input[['prs_column']])%>%mutate(data = 'prs_col')
go_combined <- data.table::rbindlist(list(rwr_row,prs_row,rwr_column,prs_column))
ci <- function(p,n){
   sd = sqrt(p*(1-p)/n)
   ci = 1.96*sd
   ci
}

prs_column$tolist  <-  str_remove_all(prs_column$study_items,'\\[|\\]')%>%
  #str_split(.,', ')%>%
  str_extract_all(.,"[^',]+[^']")# %>%list()

df <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(df) <- c("orf_name_id", "count")
for(i in unique(prs_column$orf_name_id)){
  all_set = unlist(prs_column$tolist[prs_column$orf_name_id==i])
  int_ <- length(intersect(all_set,effector_pcc$`Systematic gene name`))
  df <- bind_rows(df,data.frame(orf_name_id=i,count=int_))
}
num_of_groups_with_effectors <- df%>%filter(count>0)%>%nrow()

# New facet label names for dose variable
axis.labs <- c("Row-based", "Column-based")
names(axis.labs) <- c("row", "col")
go_summarised <- go_combined%>%
  group_by(data)%>%
  summarise(positives=n_distinct(orf_name_id))%>%
  mutate(n_tot=5183, prop = positives/n_tot, ci_min=prop-ci(prop,n_tot),ci_max=prop+ci(prop,n_tot))%>%
  separate(data,into=c('method','axis'))%>%
  mutate(axis=factor(axis,levels=c('row','col')))

p_val_row  <- go_summarised%>%
  mutate(negatives = n_tot-positives)%>%
  dplyr::filter(axis=="row")%>%
  select(positives,negatives)%>%
  as.matrix()%>%
  prop.test(correct=FALSE)
p_val_col <- go_summarised%>%
  mutate(negatives = n_tot-positives)%>%
  dplyr::filter(axis=="col")%>%
  select(positives,negatives)%>%
  as.matrix()%>%
  prop.test(correct=FALSE)

annotation_df <- data.frame(
  axis = factor(c("row", "col")),
  start = c("prs", "prs"),
  end = c("rwr", "rwr"),
  y = c(0.91, 0.91),
  label = c(glue::glue("p = {format.pval(p_val_row$p.value,digits=2)}"), glue::glue("p{format.pval(p_val_col$p.value)}"))
)

 prs_rwr_comp <- ggplot()+
  facet_grid(vars(axis),
  labeller = labeller(axis=axis.labs))+
  geom_col(data=go_summarised,aes(x=method,y=prop,fill=method))+
  geom_errorbar(data=go_summarised, aes(x=method, y=prop, ymin=ci_min, ymax=ci_max), width=.2) +
  theme_bw()+
  scale_x_discrete(labels=c('PRS','RWR'))+
  xlab('')+
  ylab('%of tests with GO enrichments')+
  theme(
   strip.background = element_rect(
     color=,"black", fill="white", size=1.5, linetype="solid"
     ),
     legend.position='none',
     panel.spacing = unit(2, "lines")
   )+
   geom_signif(
    data = annotation_df,
    aes(xmin = start, xmax = end, annotations = label, y_position = y),
    textsize = 4,tip_length = .1,
    manual = TRUE
  )+
  scale_y_continuous(expand=c(0,0),limits=c(0,1))

prs_rwr_comp
```

```{r}
if(snakemake@params['save']==TRUE){

  figure_folder = "reports/figures/paper_figures_052521/"
    ggsave(glue("{figure_folder}/fig5b.png"),plot=prs_rwr_comp, width=3,height=4)
}

```