### 2021-09-12 notebooks/06-Fig1b-091121.Rmd
converted fig1b to complexheatmap in R

---
title: "PRS dendrogram clusters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
#use_python("/home/oma21/miniconda3/envs/enm/bin/python")
use_condaenv('enm')
setwd('../')
#source('enm/R/plot_density_prs.R')
#source('enm/R/get.subclusters.dendrogram.R')
library(RColorBrewer)
col =brewer.pal(n=8,name='Set1')
sensor_colors = col[1:4]
effector_colors = col[5:8]
```

# Read data
Read data in python, from pickle file saved earlier.
```{python}
import matplotlib.pyplot as plt
import matplotlib as mpl
import networkx as nx
import numpy as np
import pandas as pd
import plotnine as p9
import pickle 
import os
import re
import itertools as itr
from enm.Enm import Enm
from enm.utils import *
import scipy.cluster.hierarchy as sch
print(os.getcwd())
with open(f'../data/interim/pcc.pickle','rb') as f:
    e_pcc = pickle.load(f)
pcc_df = e_pcc.df
pcc_gc = e_pcc.graph_gc
pcc_gc_pos = (dict(pcc_gc.nodes(data='pos')))

```

Use scipy distance calculation and pass the square form of the distance matrix to R.
This is done because R's `dist` function is significantly slower than scipy.

```{python}
prs_mat = e_pcc.prs_mat
quantile_threshold =  0.95
method = 'ward'
metric = 'seuclidean'
q99 = np.quantile(prs_mat, quantile_threshold)
prs_mat_cl = copy.deepcopy(prs_mat)
prs_mat_cl[prs_mat_cl > q99] = q99
prs_mat_cl_orig = prs_mat_cl
# , optimal_ordering=True)
#row_linkage_cl = sch.linkage(sch.distance.pdist(prs_mat_cl), method=method)
# ,optimal_ordering=True)
#col_linkage_cl = sch.linkage(sch.distance.pdist(prs_mat_cl.T), method=method)

row_dist_orig = sch.distance.pdist((prs_mat),metric = metric)
col_dist_orig = sch.distance.pdist((prs_mat.T),metric = metric)
row_dist_orig_sqr = sch.distance.squareform(row_dist_orig)
col_dist_orig_sqr = sch.distance.squareform(col_dist_orig)

row_dist = sch.distance.pdist(prs_mat_cl,metric = metric)
col_dist = sch.distance.pdist(prs_mat_cl.T,metric = metric)

row_dist_sqr = sch.distance.squareform(row_dist)
col_dist_sqr = sch.distance.squareform(col_dist)

```
# Plot heatmap
Using the clustering with `seuclidian` distance and `ward` linkage, plot the heatmap with row and column colors are the GO term annotations found from effector and sensor analysis

```{r}
library(tidyverse)
library(ComplexHeatmap)
cluster_method = 'ward.D2'
dist_row = as.dist(py$row_dist_sqr)
hclust_row = hclust(dist_row, cluster_method)
dist_col = as.dist(py$col_dist_sqr)
hclust_col = hclust(dist_col, cluster_method)
row_dend = as.dendrogram(hclust_row)%>% dendextend::color_branches(k=2, col = c('black', '#f15a24'))
col_dend = as.dendrogram(hclust_col)%>%  dendextend::color_branches(k=2, col = c('#f7931e', 'black'))
#row_dend <- row_dend %>%

# go_color_map = unlist(append(py$effector_colors_dict,py$sensor_colors_dict))
# go_color_map['none'] = 'white'
# clrs = go_color_map# map(py$eff_sens_go_list, ~ifelse(is.null(.x), 'none',.x))%>%unlist()

# rowColors = rowAnnotation(rowColors=py$e_pcc$df%>%left_join(py$effector_pcc)%>%select(go_term)%>%pull%>%map(~ifelse(is.na(.x), 'none',.x))%>%unlist(),
#                           col=list(rowColors = go_color_map ),
#                           #width=unit(100,'mm'),
#                           simple_anno_size=unit(3,'cm'),
#                           show_annotation_name=FALSE,
#                           show_legend=FALSE)
# colColors = columnAnnotation(colColors=py$e_pcc$df%>%left_join(py$sensors_pcc)%>%select(go_term)%>%pull%>%map(~ifelse(is.null(.x), 'none',
#                                                                                                                       ifelse(is.nan(.x),'none',.x)))%>%unlist(),
#                           col=list(colColors = go_color_map ),
#                           simple_anno_size=unit(3,'cm'),
#                           #annotation_height=unit(20,'mm'),
#                           show_annotation_name=FALSE,
#                           show_legend=FALSE)
rowAnnot = HeatmapAnnotation(Effectiveness=anno_lines(rowSums(py$prs_mat), border=FALSE,
                            width = unit(0.2, 'in'),
                            axis=FALSE),
                            which='row',
                            #name='', 
                            gp = gpar(col='blue'),
                            annotation_name_side = 'top',
                            annotation_name_rot= 90,
                            border=FALSE,
                            #annotation_name_gp= gpar(fontsize = 0)
                            show_annotation_name=FALSE,
                          show_legend=FALSE
)
colAnnot = HeatmapAnnotation(Sensitivity=anno_lines(colSums(py$prs_mat), 
                            axis = FALSE, border=FALSE,height = unit(0.2, 'in')), 
                             which='col', 
                            # name='',
                            gp = gpar(col='blue'),
                            #annotation_name_gp= gpar(fontsize = 0)
                            annotation_name_side = 'left',
                            annotation_name_rot=0,
                            border=FALSE,
                            show_annotation_name=FALSE,
                          show_legend=FALSE
)

png('tmp_heatmap.png', width = 3, height=2.5, units='in', res=150 )
hm<-Heatmap(py$prs_mat_cl, 
        cluster_rows = row_dend,#as.dendrogram(hclust_row),
        cluster_columns= col_dend,#as.dendrogram(hclust_col),
        col=RColorBrewer::brewer.pal(9,'YlGnBu'),
        row_dend_width = unit(0.2,'in'),
        column_dend_height = unit(0.2,'in'),
        right_annotation=rowAnnot,
#        left_annotation = rowColors,
        bottom_annotation=colAnnot,
#        top_annotation=colColors,
show_heatmap_legend = FALSE,
)
draw(hm ,padding = unit(c(0, 0.5, 0, 0), "in"))

dev.off()

png('tmp_heatmap_lgn.png', width = 3, height=2.5, units='in', res=150 )

lgd = Legend(title = "", 
            at = c(0, 1e-4,2e-4,3e-4,4e-4), col_fun = circlize::colorRamp2(c(0, 1e-4,2e-4,3e-4,4e-4), RColorBrewer::brewer.pal(5,'YlGnBu')) , labels_gp = gpar(font=6), title_position = 'leftcenter-rot', legend_width=unit(0.05, 'in'),labels_rot=45)
#
draw(lgd, x = unit(0.1, "npc"), y = unit(0.8, "npc"))

dev.off()
```